---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---

```{r}
library(dplyr)
library(Seurat)
library(gProfileR)
```


```{r}
all10x <- readRDS('output/10x-180504')
```

One cluster in the data contains cells from all samples

```{r}
TSNEPlot(all10x, group.by='sample_name', pt.size=0.1)
```

Cluster 12

```{r}
TSNEPlot(all10x, group.by='res.0.5', pt.size=0.1, do.label=T)
```

#Marker genes

```{r}
markers <- read.table('output/markergenes/180504/markers_10x-180504_res.0.5_negbinom', sep='\t', header=T)
mixture <- markers[markers$cluster == 12,]
mixture <- mixture[mixture$p_val_adj < 0.05,]
mixture_pos <- mixture[order(-mixture$avg_logFC),]
mixture_neg <- mixture[order(mixture$avg_logFC),]
```

Positive markers 

```{r}
as.data.frame(mixture_pos)
```

Negative markers

```{r}
as.data.frame(mixture_neg)
```

Top 10 positive markers 

```{r fig1, fig.height = 15, fig.width = 10, fig.align = "center"}
VlnPlot(all10x, features.plot=toupper(mixture_pos$gene[1:10]), group.by='res.0.5', point.size.use=-1, nCol=2, x.lab.rot=T, size.x.use=10)
```

Top 10 negative markers

```{r fig2, fig.height = 15, fig.width = 10, fig.align = "center"}
VlnPlot(all10x, features.plot=toupper(mixture_neg$gene[1:10]), group.by='res.0.5', point.size.use=-1, nCol=2, x.lab.rot=T, size.x.use=10)
```

Plots

```{r fig4, fig.height = 5, fig.width = 9, fig.align = "center"}
all10x@meta.data$mixture <- ifelse(all10x@meta.data$res.0.5==12, "mixture", "rest")
VlnPlot(all10x, features.plot=c('nGene', 'MALAT1', 'NEAT1', 'FN1', 'ITGB1', 'COL1A1', 'COL1A2', 'percent.mito'), group.by='mixture', point.size.use=-1, nCol=4)
```


#Gene set enrichment

```{r}
gsea_revigo <- read.table('tables/Revigo_mixture-cluster_cleaned.txt', header=T, sep='\t')
gsea_revigo
```


#Nr of cells

Sample composition in cluster 12. 

```{r}
cluster12 <- SubsetData(all10x, cells.use=rownames(all10x@meta.data)[which(all10x@meta.data$res.0.5 %in% 12)])
rotate_x <- function(data, column_to_plot, labels_vec, rot_angle) {
     plt <- barplot(data[[column_to_plot]], col='steelblue', xaxt="n")
     text(plt, par("usr")[3], labels = labels_vec, srt = rot_angle, adj = c(1.1,1.1), xpd = TRUE, cex=1)
}
rotate_x((cluster12@meta.data %>% count(sample_name))[,2], 'n', as.vector(unlist((cluster12@meta.data %>% count(sample_name))[,1])), 45)
```

#scmap results

```{r}
TSNEPlot(all10x, group.by='predicted_labels_fat', pt.size=0.1)
```

Add line breaks for legend

```{r}
all10x@meta.data$predicted_labels_fat_breaks = as.vector(all10x@meta.data$predicted_labels_fat)
all10x@meta.data$predicted_labels_fat_breaks[which(all10x@meta.data$predicted_labels_fat == 'mesenchymal stem cell of adipose')] = 'mesenchymal\nstem cell\nof adipose'
all10x@meta.data$predicted_labels_fat_breaks[which(all10x@meta.data$predicted_labels_fat == 'smooth muscle cell')] = 'smooth\nmuscle cell'
```

```{r}
TSNEPlot(all10x, group.by='predicted_labels_fat_breaks', pt.size=0.1) + theme(legend.key.height=unit(1, "cm"))
```


#Stressed cell analysis

```{r}
stress_genes <- read.table('/raid5/projects/timshel/sc-arc_lira/src/data-genelists/171219-van_den_Brink2017-genes_affected_by_dissociation.csv', header=T) %>% pull(1)
length(stress_genes)
```

```{r}
stress_genes
```


```{r}
de_genes <- read.table('output/markergenes/180504/markers_10x-180504_res.0.5_negbinom', header=T)
de_genes <- de_genes[de_genes$p_val_adj < 0.05,]
```

Cluster 12 is the mixture cluster. Check markers MALAT1 and NEAT1 to be sure:

```{r fig7, fig.height = 3, fig.width = 10, fig.align = "center"}
VlnPlot(all10x, group.by='res.0.5', features.plot=c('MALAT1', 'NEAT1'), point.size.use =-1)
```

Are any of the stress genes in the DE genes for the mixture cluster?

```{r}
de_genes_pos <- de_genes[de_genes$avg_logFC > 0, ]
de_genes_neg <- de_genes[de_genes$avg_logFC < 0, ]

intersect_pos <- list()
intersect_neg <- list()
n_pos <- list()
n_neg <- list()
perc_pos <- list()
perc_neg <- list()

for (i in unique(de_genes$cluster)){
  c <- paste('cluster', i)

  genes_pos <- de_genes_pos[de_genes_pos$cluster == i, 'gene']
  genes_neg <- de_genes_neg[de_genes_neg$cluster == i, 'gene']
  
  intersect_pos[[c]] <- length(intersect(toupper(stress_genes), genes_pos))
  intersect_neg[[c]] <- length(intersect(toupper(stress_genes), genes_neg))
  n_pos[[c]] <- length(genes_pos)
  n_neg[[c]] <- length(genes_neg)
  perc_pos[[c]] <- (intersect_pos[[c]] / n_pos[[c]]) * 100
  perc_neg[[c]] <- (intersect_neg[[c]] / n_neg[[c]]) * 100
}

stressed_cells <- data.frame(
  shared.pos.genes=unlist(intersect_pos),
  shared.neg.genes=unlist(intersect_neg),
  n_pos=unlist(n_pos),
  n_neg=unlist(n_neg),
  perc_pos=unlist(perc_pos),
  perc_neg=unlist(perc_neg)
  )

stressed_cells <- stressed_cells[order(rownames(stressed_cells)),]
stressed_cells
```

3 of the 140 genes were found in the positive markers for the mixture cluster, 54 in the negative numbers. Similar numbers were found for the other clusters in the data. These results indicate that the mixture cluster does not consist of stressed cells. 

```{r}

```


#Figures for report

```{r fig5, fig.height = 9, fig.width = 12, fig.align = "center"}
get_violin_plot <- function(x){
  return(VlnPlot(all10x, features.plot=c(x), point.size.use=-1, group.by='mixture', size.title.use=14, remove.legend=T) + theme(axis.title.x=element_blank()))
}

fig2_first_row <- plot_grid(
  TSNEPlot(all10x, group.by='sample_name', pt.size=0.1),
  TSNEPlot(all10x, group.by='predicted_labels_fat_breaks', pt.size=0.1) + theme(legend.key.height=unit(1, "cm")),
  labels=c('a', 'b'), rel_widths=c(48/100, 52/100)
)

fig2_violin <- plot_grid(
  get_violin_plot('nGene'),
  get_violin_plot('MALAT1'),
  get_violin_plot('NEAT1'),
  get_violin_plot('FN1'),
  get_violin_plot('ITGB1'),
  get_violin_plot('COL1A1'),
  get_violin_plot('COL1A2'),
  get_violin_plot('percent.mito'),
  labels=c('d', 'e', 'f', 'g', 'h', 'i', 'j', 'k'), nrow=2
)

fig2_barplot <- plot_grid(ggplot(data=cluster12@meta.data %>% count(sample_name), aes(x=sample_name, y=n)) +
  geom_bar(stat="identity", position='dodge') + 
  scale_y_continuous(expand=c(0,0)) +
  theme(axis.line.y=element_blank(), axis.ticks.y=element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank()) +
  coord_flip(), labels=c('c'))

fig2_second_row <- plot_grid(fig2_barplot, fig2_violin, rel_widths=c(2/8, 6/8))

fig2 <- plot_grid(
  fig2_first_row,
  fig2_second_row,
  nrow=2
)

fig2

save_plot("plots/180504_mixture.pdf", fig2, base_width=12, base_height=9)

```


```{r fig6, fig.height = 10, fig.width = 12, fig.align = "center"}
# get_violin_plot <- function(x){
#   return(VlnPlot(all10x, features.plot=c(x), point.size.use=-1, group.by='mixture', size.title.use=14, remove.legend=T) + theme(axis.title.x=element_blank()))
# }
# 
# fig2_first_row <- plot_grid(
#   TSNEPlot(all10x, group.by='sample_name', pt.size=0.1) + theme(legend.position=c(0.1, 0.5), legend.key = element_blank(), legend.background= element_rect(fill=alpha('white', 0.8), size=0.5, linetype='solid', colour=alpha('black', 0.3))),
#   DimPlot(all10x, reduction.use='tsne', cells.highlight = rownames(all10x@meta.data)[all10x@meta.data$res.0.5 == 12], cols.highlight='blue', cols.use='gray', pt.size=0.1),
#   labels=c('a', 'b')
# )
# 
# fig2_violin <- plot_grid(
#   get_violin_plot('nGene'),
#   get_violin_plot('MALAT1'),
#   get_violin_plot('NEAT1'),
#   get_violin_plot('FN1'),
#   get_violin_plot('ITGB1'),
#   get_violin_plot('COL1A1'),
#   get_violin_plot('COL1A2'),
#   get_violin_plot('percent.mito'),
#   labels=c('d', 'e', 'f', 'g', 'h', 'i', 'j', 'k'), nrow=2
# )
# 
# fig2_barplot <- plot_grid(ggplot(data=cluster12@meta.data %>% count(sample_name), aes(x=sample_name, y=n)) +
#   geom_bar(stat="identity", position='dodge') + 
#   scale_y_continuous(expand=c(0,0)) +
#   theme(axis.line.y=element_blank(), axis.ticks.y=element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank()) +
#   coord_flip(), labels=c('c'))
# 
# fig2_second_row <- plot_grid(fig2_barplot, fig2_violin, rel_widths=c(2/8, 6/8))
# 
# fig2 <- plot_grid(
#   fig2_first_row,
#   fig2_second_row,
#   nrow=2
# )
# 
# fig2
# 
# #save_plot("plots/180504_mixture.pdf", fig2, base_width=12, base_height=10)

```


```{r}
write.table(stressed_cells, 'tables/10x-180504-stressed-cells-analysis.txt', sep='\t')
```

