---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---


```{r}
library(Seurat)
library(dplyr)
source('code/colors.R')
seurobj <- readRDS('output/seurat_objects/180831/10x-180831')
```

#Quality control plots

The number of genes expressed descreases over time.

```{r fig1, fig.height=15, fig.width=10, fig.align="center"}
VlnPlot(seurobj, c("nGene", "percent.mito", "nUMI"), group.by='timepoint', nCol = 1, point.size.use=-1, size.x.use = 10)
```

```{r}
GenePlot(seurobj, 'nUMI', 'nGene', cex.use = 0.5)
```

```{r}
PCElbowPlot(seurobj, num.pc=50) #TSNE+clustering run on 21 PC's.
```

#PCA

Interesting to see: T4 and T5 contain a lot more variation than T1, T2 and T3, and PC2 seems to split T4 and T5. Could the split in PC2 describe the cells developing into white or brown? 

```{r}
PCAPlot(seurobj, group.by='timepoint', pt.size=0.1)
```

A few clusters in the data have much higher expression of 'ADIPOQ', 'SCD', 'RBP4', 'G0S2', 'PLIN4', 'FABP5'. This seems to be captured by PC2. 

```{r fig2, fig.height = 10, fig.width = 10, fig.align = "center"}
FeaturePlot(seurobj, reduction.use='pca', features.plot=c('ADIPOQ', 'SCD', 'RBP4', 'G0S2', 'PLIN4', 'FABP5'), pt.size=1, cols.use=c('gray', 'blue'), no.legend=F, nCol=2)
```



```{r fig3, fig.height = 10, fig.width = 10, fig.align = "center"}
FeaturePlot(seurobj, reduction.use='pca', features.plot=c('PLA2G2A', 'MT1X', 'APOD', 'DPT', 'PTGDS', 'IGF2'), pt.size=1, cols.use=c('gray', 'blue'), no.legend=F, nCol=2)
```

PLA2G2A: http://www.jlr.org/content/early/2017/06/29/jlr.M076141 
"...suggesting that PLA2G2A activates mitochondrial uncoupling in brown adipose tissue."


PDGFRα/PDGFRβ signaling balance modulates progenitor cell differentiation into white and beige adipocytes. Based on PDGFRα or PDGFRβ deletion and ectopic expression experiments, we conclude that the PDGFRα/PDGFRβ signaling balance determines progenitor commitment to beige (PDGFRα) or white (PDGFRβ) adipogenesis. Our study suggests that adipocyte lineage specification and metabolism can be modulated through PDGFR signaling. http://dev.biologists.org/content/145/1/dev155861.long 

```{r fig4, fig.height = 3, fig.width = 10, fig.align = "center"}
FeaturePlot(seurobj, reduction.use='pca', features.plot=c('PDGFRA', 'PDGFRB'), pt.size=1, cols.use=c('gray', 'blue'), no.legend=F, nCol=2)
```

#TSNE

```{r}
TSNEPlot(seurobj, group.by='timepoint', pt.size=0.1)
```


```{r}
TSNEPlot(seurobj, group.by='Phase', pt.size=0.1)
```

Cluster 11 = mixture cluster.

```{r}
TSNEPlot(seurobj, group.by='res.0.5', pt.size=0.1, do.label=T)
```

```{r fig5, fig.height = 3, fig.width = 10, fig.align = "center"}
VlnPlot(seurobj, group.by='res.0.5', features.plot=c('MALAT1', 'NEAT1'), point.size.use=-1)
```

#Metadata feature plots

```{r}
FeaturePlot(seurobj, reduction.use='tsne', features.plot = 'nUMI', cols.use=c('grey', 'blue'), no.legend=F)
```

```{r}
FeaturePlot(seurobj, features.plot = 'percent.mito', cols.use=c('grey', 'blue'), no.legend = F)
```

```{r}
FeaturePlot(seurobj, reduction.use='tsne', features.plot = 'nGene', cols.use=c('grey', 'blue'), no.legend = F)
```

#Markergene expression

```{r}
FeaturePlot(seurobj, reduction.use='tsne', features.plot = 'EBF2', cols.use=c('grey', 'blue'), no.legend = F)
```

```{r}
FeaturePlot(seurobj, reduction.use='tsne', features.plot = 'TM4SF1', cols.use=c('grey', 'blue'), no.legend = F)
```

```{r}
FeaturePlot(seurobj, reduction.use='tsne', features.plot = 'LY6K', cols.use=c('grey', 'blue'), no.legend = F)
```

```{r}
FeaturePlot(seurobj, reduction.use='tsne', features.plot = 'PDGFRA', cols.use=c('grey', 'blue'), no.legend = F)
```


#Markers literature

Marker genes for mature brown/beige compared to white mentioned by Seale 2016: UCP1, DIO2, CIDEA, PPARGC1A, PPARA, COX7A1, COX8B, PRDM16, EBF2. \\

```{r fig6, fig.height = 12, fig.width = 10, fig.align = "center"}
VlnPlot(seurobj, features.plot=c('UCP1', 'DIO2', 'CIDEA', 'PPARGC1A', 'PPARA', 'COX7A1', 'PRDM16', 'EBF2'), group.by='timepoint', point.size.use = -1, nCol=2)
```


Based on PDGFRα or PDGFRβ deletion and ectopic expression experiments, we conclude that the PDGFRα/PDGFRβ signaling balance determines progenitor commitment to beige (PDGFRα) or white (PDGFRβ) adipogenesis. Our study suggests that adipocyte lineage specification and metabolism can be modulated through PDGFR signaling. http://dev.biologists.org/content/145/1/dev155861.long 

```{r fig7, fig.height = 3, fig.width = 10, fig.align = "center"}
FeaturePlot(seurobj, reduction.use='pca', features.plot=c('PDGFRA', 'PDGFRB'), pt.size=1, cols.use=c('gray', 'blue'), no.legend=F, nCol=2)
```

```{r fig8, fig.height = 3, fig.width = 10, fig.align = "center"}
FeaturePlot(seurobj, reduction.use='tsne', features.plot=c('PDGFRA', 'PDGFRB'), pt.size=1, cols.use=c('gray', 'blue'), no.legend=F, nCol=2)
```

```{r}
GenePlot(SetAllIdent(seurobj, id='timepoint'), gene1='PDGFRA', gene2='PDGFRB', cex.use=0.5)
```

#Figures for report

```{r fig9, fig.height = 5, fig.width = 12, fig.align = "center"}
fig <- plot_grid(
  PCAPlot(seurobj, group.by='timepoint', pt.size=0.1),
  TSNEPlot(seurobj, group.by='timepoint', pt.size=0.1),
  labels='auto', nrow=1
)

#save_plot('../figures/figures_paper/supplementary_figures/pca/180831_pca_tsne.pdf', PCAPlot(seurobj, group.by='timepoint', cols.use=colors.timepoints, pt.size=0.1, do.return=T), base_width=6, base_height=5)
#save_plot('../figures/figures_paper/supplementary_figures/pca/180831_pca_tsne.png', PCAPlot(seurobj, group.by='timepoint', cols.use=colors.timepoints, pt.size=0.1, do.return=T), base_width=6, base_height=5)

#save_plot("../plots/180831_pca_tsne.pdf", fig, base_width=12, base_height=5)
fig
```

```{r fig10, fig.height = 4, fig.width = 6, fig.align = "center"}
sfig1 <- PCElbowPlot(seurobj, num.pc=50) #TSNE+clustering run on 21 PC's.
#save_plot("../plots/supplementary_figures/sfig_180831_pcelbow.pdf", sfig1, base_width=6, base_height=4)
sfig1

```

```{r fig11, fig.height = 4.5, fig.width = 6, fig.align = "center"}
sfig2 <- TSNEPlot(seurobj, group.by='Phase', pt.size=0.1)
#save_plot("../plots/supplementary_figures/sfig_180831_tsne_cellcycle.pdf", sfig2, base_width=6, base_height=4.5)
sfig2
```

```{r fig12, fig.height = 3, fig.width = 12, fig.align = "center"}
sfig3 <- plot_grid(
  VlnPlot(seurobj, c("nGene"), group.by='timepoint', point.size.use=-1, cols.use=colors.timepoints),
  VlnPlot(seurobj, c("nUMI"), group.by='timepoint', point.size.use=-1, cols.use=colors.timepoints),
  VlnPlot(seurobj, c("percent.mito"), group.by='timepoint', point.size.use=-1, cols.use=colors.timepoints),
  labels='auto', nrow=1
)
#save_plot("../figures/figures_paper/supplementary_figures/ngene-numi-pm/sfig_180831_ngene-numi-pm.pdf", sfig3, base_width=12, base_height=3)
#save_plot("../figures/figures_paper/supplementary_figures/ngene-numi-pm/sfig_180831_ngene-numi-pm.png", sfig3, base_width=12, base_height=3)
sfig3
```

```{r fig13, fig.height = 9, fig.width = 12, fig.align = "center"}
sfig4 <- plot_grid(
  TSNEPlot(seurobj, group.by='res.0.5', do.label=T, pt.size=0.1),
  TSNEPlot(seurobj, group.by='res.0.7', do.label=T, pt.size=0.1),
  TSNEPlot(seurobj, group.by='res.1', do.label=T, pt.size=0.1),
  TSNEPlot(seurobj, group.by='res.1.5', do.label=T, pt.size=0.1),
  labels='auto', nrow=2
)
sfig4
#save_plot("../plots/supplementary_figures/sfig_180831_clusters.pdf", sfig4, base_width=12, base_height=9)
```

```{r}
#T1T2T3 <- readRDS('output/10x-180831-T1T2T3')
#T4T5 <- readRDS('output/10x-180831-T4T5')
```

```{r fig14, fig.height = 9, fig.width = 12, fig.align = "center"}
#sfig5 <- plot_grid(
#  TSNEPlot(T1T2T3, group.by='timepoint', pt.size=0.1, colors.use=c('#f8766d', '#a3a500', #'#00bf7d')),
#  TSNEPlot(T4T5, group.by='timepoint', pt.size=0.1, colors.use=c('#00b0f6', '#e76bf3')),
#  TSNEPlot(T1T2T3, group.by='res.1.5', pt.size=0.1, do.label=T),
#  TSNEPlot(T4T5, group.by='res.1.5', pt.size=0.1, do.label=T),
#  labels=c('b', 'c', 'd', 'e')
#)
#save_plot("../plots/supplementary_figures/sfig_180831_feature_selection.pdf", sfig5, base_width=12, base_height=9)
#sfig5
```

