---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Notebook for Monocle trajectories of the 180504 dataset.

To investigate the robustness of the trajectories, the dataset was randomly downsampled on 1000 cells per sample. 4 different types of regressions were used when running the DDRTree algorithm.


```{r results='hide', message=FALSE, warning=FALSE}
library(Seurat)
library(monocle)
```

Loading all datasets.

```{r}
cds.11 <- readRDS('output/monocle/180504/10x-180504-monocle-11')
cds.cc.11 <- readRDS('output/monocle/180504/10x-180504-monocle-cc-11')
cds.pm.umi.11 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-11')
cds.pm.umi.cc.11 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-cc-11')

cds.27 <- readRDS('output/monocle/180504/10x-180504-monocle-27')
cds.cc.27 <- readRDS('output/monocle/180504/10x-180504-monocle-cc-27')
cds.pm.umi.27 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-27')
cds.pm.umi.cc.27 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-cc-27')

cds.33 <- readRDS('output/monocle/180504/10x-180504-monocle-33')
cds.cc.33 <- readRDS('output/monocle/180504/10x-180504-monocle-cc-33')
cds.pm.umi.33 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-33')
cds.pm.umi.cc.33 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-cc-33')

cds.53 <- readRDS('output/monocle/180504/10x-180504-monocle-53')
cds.cc.53 <- readRDS('output/monocle/180504/10x-180504-monocle-cc-53')
cds.pm.umi.53 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-53')
cds.pm.umi.cc.53 <- readRDS('output/monocle/180504/10x-180504-monocle-pm-umi-cc-53')
```

#Trajectories of one subset

Trajectories of one subset, coloured by sample and cell cycle phase. Topleft: no regression. Topright: cell cycle regression. Bottom left: percent.mito + nUMI regression. Bottom right: percent.mito + nUMI + cell cycle regression.

```{r fig1, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.11, color_by='depot'),
  plot_cell_trajectory(cds.cc.11, color_by='depot'),
  plot_cell_trajectory(cds.pm.umi.11, color_by='depot'),
  plot_cell_trajectory(cds.pm.umi.cc.11, color_by='depot'),
  nrow=2)
```

```{r fig2, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.11, color_by='Phase'),
  plot_cell_trajectory(cds.cc.11, color_by='Phase'),
  plot_cell_trajectory(cds.pm.umi.11, color_by='Phase'),
  plot_cell_trajectory(cds.pm.umi.cc.11, color_by='Phase'),
  nrow=2)
```


#Trajectories, no regression

Trajectories of the 4 subsets with no variables regressed out. The ordering of the cells seems to be influenced by the cell cycle state (though not completely). Interestingly in three of the trajectories there is a branch split. 

```{r fig3, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.11, color_by='Phase'),
  plot_cell_trajectory(cds.27, color_by='Phase'),
  plot_cell_trajectory(cds.33, color_by='Phase'),
  plot_cell_trajectory(cds.53, color_by='Phase'),
  nrow=2)
```

Does the branch split discriminate between depots?

```{r fig4, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.11, color_by='depot'),
  plot_cell_trajectory(cds.27, color_by='depot'),
  plot_cell_trajectory(cds.33, color_by='depot'),
  plot_cell_trajectory(cds.53, color_by='depot'),
  nrow=2)
```

#Trajectories, cell cycle regression

Trajectories of the 4 subsets with cell cycle effects regressed out. Topleft, bottom left and bottom right look similar but all have different number of branching points. 

```{r fig5, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.cc.11, color_by='Phase'),
  plot_cell_trajectory(cds.cc.27, color_by='Phase'),
  plot_cell_trajectory(cds.cc.33, color_by='Phase'),
  plot_cell_trajectory(cds.cc.53, color_by='Phase'),
  nrow=2)
```

Coloured on depot. Branches do not 

```{r fig6, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.cc.11, color_by='depot'),
  plot_cell_trajectory(cds.cc.27, color_by='depot'),
  plot_cell_trajectory(cds.cc.33, color_by='depot'),
  plot_cell_trajectory(cds.cc.53, color_by='depot'),
  nrow=2)
```

#Trajectories, percent.mito + nUMI regression

In general all cells seem to follow one direction, and the general shape of the trajectories are similar. The branch splits in the trajectories above (no regression and cell cycle regression) could be the effects of differences in percent.mito and nUMI, and that's why we don't see that here. 

```{r fig7, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.pm.umi.11, color_by='Phase'),
  plot_cell_trajectory(cds.pm.umi.27, color_by='Phase'),
  plot_cell_trajectory(cds.pm.umi.33, color_by='Phase'),
  plot_cell_trajectory(cds.pm.umi.53, color_by='Phase'),
  nrow=2)
```

#Trajectories, percent.mito + nUMI + cell cycle regression

Trajectories with percent.mito, nUMI and cell cycle effects regressed out. Doesn't look very convincing. 

```{r fig8, fig.height = 8, fig.width = 10, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(cds.pm.umi.cc.11, color_by='depot'),
  plot_cell_trajectory(cds.pm.umi.cc.27, color_by='depot'),
  plot_cell_trajectory(cds.pm.umi.cc.33, color_by='depot'),
  plot_cell_trajectory(cds.pm.umi.cc.53, color_by='depot'),
  nrow=2)
```

#Figures for report

a = trajectory without regressions, coloured on cell cycle phase. b = trajectory with cell cycle effects regressed out, coloured on depot. 

```{r fig9, fig.height = 5, fig.width = 12, fig.align = "center"}
fig <- plot_grid(
  plot_cell_trajectory(cds.11, color_by='Phase'),
  plot_cell_trajectory(cds.cc.11, color_by='depot'),
  nrow=1, labels='auto')
save_plot("plots/180504_monocle_noreg-ccreg.pdf", fig, base_width=12, base_height=5)
fig
```

Supplementary figures

```{r fig10, fig.height = 12, fig.width = 12, fig.align = "center"}
noreg <- plot_grid(
  plot_cell_trajectory(cds.11, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.27, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.33, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.53, color_by='Phase', cell_size = 0.5),
  nrow=1)

ccreg <- plot_grid(
  plot_cell_trajectory(cds.cc.11, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.cc.27, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.cc.33, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.cc.53, color_by='Phase', cell_size = 0.5),
  nrow=1)

pmumireg <- plot_grid(
  plot_cell_trajectory(cds.pm.umi.11, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.pm.umi.27, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.pm.umi.33, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.pm.umi.53, color_by='Phase', cell_size = 0.5),
  nrow=1)

pmumiccreg <- plot_grid(
  plot_cell_trajectory(cds.pm.umi.cc.11, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.pm.umi.cc.27, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.pm.umi.cc.33, color_by='Phase', cell_size = 0.5),
  plot_cell_trajectory(cds.pm.umi.cc.53, color_by='Phase', cell_size = 0.5),
  nrow=1)

sfig <- plot_grid(
  noreg,
  ccreg,
  pmumireg,
  pmumiccreg,
  labels='auto', nrow=4
)

save_plot("plots/supplementary_figures/sfig_180504_monocle.pdf", sfig, base_width=12, base_height=12)
sfig
```

