---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---


```{r}
library(GOplot)
library(ggplot2)
library(ggrepel)
```



```{r}
name <- 'heatmap_logFC0.3_ncluster6'
folder <- paste('../tables/tables_paper/all_tables/BEAM/', name, sep='')

clusters <- c(1,2,3,4,6)

gsea <- list()
genelist <- list()

for (i in clusters){
  gsea_cluster <- read.table(paste(folder, '/REVIGO_cleaned/', 'REVIGO_cluster', i, '_cleaned.tsv', sep=''), header=T)
  gsea_cluster['cluster'] <- i
  genes_cluster <- read.table(paste(folder, '/genelist_cluster', i, '.tsv', sep=''), header=T)
  gsea[[i]] <- gsea_cluster
  genelist[[i]] <- genes_cluster
}

gsea <- do.call("rbind", gsea)
genelist <- do.call("rbind", genelist)
```

```{r}
create_terms <- function(gsea){
  return(data.frame(category=gsea$gsea.domain, ID=gsea$revigo.term_ID, term=gsea$revigo.description, adj_pval=gsea$gsea.p.value, genes=gsea$gsea.intersection, cluster=gsea$cluster))
}

create_genes <- function(genelist){
  return(data.frame(ID=genelist$gene_short_name, logFC=genelist$avgLogFC_State2_State3))
}
```

#All terms

```{r}
terms <- create_terms(gsea)
genes <- create_genes(genelist)
circ <- circle_dat(terms, genes)
```

```{r, fig.height = 10, fig.width = 20, fig.align = "center"}
GOBar(circ)
```

```{r, fig.height = 12, fig.width = 25, fig.align = "center"}
GOBubble(circ, labels=8, ID=F)
```

```{r}
GOBubble_modified <- function(data, display, title, colour, labels, ID = T, table.legend = T, table.col = T, bg.col = F, zscore_threshold=0, ggrepel_linecol='black'){
  zscore <- adj_pval <- category <- count <- id <- term <- NULL
  if (missing(display)) display <- 'single'
  if (missing(title)) title <- ''
  if (missing(colour)) cols <- c("chartreuse4", "brown2", "cornflowerblue") else cols <- colour
  if (missing(labels)) labels <- 5
  if (bg.col == T & display == 'single') cat("Parameter bg.col will be ignored. To use the parameter change display to 'multiple'")
  
  colnames(data) <- tolower(colnames(data))
  tmp_lab <- data.frame(cats = c('Biological Process', 'Cellular Component', 'Molecular Function'), 
                        cat.abb = c('BP', 'CC', 'MF'),
                        locy = c(match('BP', data$category), match('CC', data$category), match('MF', data$category)),
                        cols = cols,
                        stringsAsFactors = F)
  tmp_lab <- tmp_lab[order(tmp_lab$locy),]
  if(!'count'%in%colnames(data)){
    rang <- c(5, 5)
    data$count <- rep(1, dim(data)[1])
  }else {rang <- c(1, 30)}
  data$adj_pval <- -log(data$adj_pval, 10)
  sub <- data[!duplicated(data$term), ]
  g <- ggplot(sub, aes(zscore, adj_pval, fill = category, size = count))+
    labs(title = title, x = 'z-score', y = '-log (adj p-value)')+
    geom_point(shape = 21, col = 'black', alpha = 1 / 2)+ #add all bubbles
    geom_hline(yintercept = 1.3, col = 'orange')+ #add threshold for p-value=0.05
    scale_size(range = rang, guide = 'none')
  
  sub2 <- subset(sub, subset=sub$adj_pval >= labels)
  sub2 <- subset(sub2, subset=abs(sub2$zscore) >= zscore_threshold)
  
  #if (!is.character(labels)) sub2 <- subset(sub, subset = sub$adj_pval >= labels) else sub2 <- subset(sub, sub$id%in%labels | sub$term%in%labels)
  if (display == 'single'){
    g <- g + scale_fill_manual('', values = tmp_lab$cols, labels = tmp_lab$cats)+
      theme(legend.position = 'bottom')+
      annotate ("text", x = 0, y = 1.8, label = "Threshold", colour = "orange", size = 4)
    if (ID) g <- g+ geom_text_repel(data = sub2, aes(x = zscore, y = adj_pval, label = id), size = 5) else g <- g + geom_text_repel(data = sub2, aes(x = zscore, y = adj_pval, label = term), size = 4, segment.color = ggrepel_linecol)
    if (table.legend){
      if (table.col) table <- draw_table(sub2, par.order = tmp_lab) else table <- draw_table(sub2)
      g <- g + theme(axis.text = element_text(size = 14), axis.line = element_line(colour = 'grey80'), axis.ticks = element_line(colour = 'grey80'), 
                     axis.title = element_text(size = 14, face = 'bold'), panel.background = element_blank(), panel.grid.minor = element_blank(), 
                     panel.grid.major = element_line(colour = 'grey80'), plot.background = element_blank()) 
      graphics::par(mar = c(0.1, 0.1, 0.1, 0.1))
      grid.arrange(g, table, ncol = 2)
    }else{
      g + theme(axis.text = element_text(size = 14), axis.line = element_line(colour = 'grey80'), axis.ticks = element_line(colour = 'grey80'), 
                axis.title = element_text(size = 14, face = 'bold'), panel.background = element_blank(), panel.grid.minor = element_blank(), 
                panel.grid.major = element_blank(), plot.background = element_blank())
    }
  }
}
```

```{r, fig.height = 9, fig.width = 15, fig.align = "center"}
p <- GOBubble_modified(circ, labels=3, ID=F, table.legend = F, zscore_threshold=4.5, ggrepel_linecol='transparent')
p
```

```{r}
save_plot("../figures/figures_paper/all_figures/plots_BEAM_GO/180831_beamGO_all-terms.pd_labels3_zscore4.5.pdf", p, base_width=18, base_height=9)
save_plot("../figures/figures_paper/main_figures/figure3/180831_beamGO_all-terms.pd_labels3_zscore4.5.pdf", p, base_width=18, base_height=9)
```


#Terms associated with metabolic branch

Cluster 2 and 6 (up in metabolic branch) and cluster 4 (down in metabolic branch).

```{r}
gsea_filtered_metabolic <- gsea[gsea$cluster %in% c(2,6,4),]
terms_metabolic <- create_terms(gsea_filtered_metabolic)
#genes <- rbind(genes_c2_swapped, genes_c5_swapped, genes_c4_swapped)
circ_metabolic <- circle_dat(terms_metabolic, genes)
```


```{r, fig.height = 10, fig.width = 20, fig.align = "center"}
GOBar(circ_metabolic)
```

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_metabolic, labels=8, ID=F)
```

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_metabolic, labels=8, ID=F, display='mutliple')
```

#Terms associated with ECM branch

Cluster 1 (down in ECM branch) and cluster 3 (up in ECM branch)

```{r}
gsea_filtered_ecm <- gsea[gsea$cluster %in% c(1,3),]
terms_ecm <- create_terms(gsea_filtered_ecm)
genes_swapped <- data.frame(ID=genes$ID, logFC=genes$logFC*-1) #swap fold change direction
circ_ecm <- circle_dat(terms_ecm, genes_swapped)
```

```{r, fig.height = 10, fig.width = 20, fig.align = "center"}
GOBar(circ_ecm)
```

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_ecm, labels=2, ID=F, table.legend=T)
```

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_ecm, labels=2, ID=F, display='mutliple')
```

#Only upregulated terms

```{r}
gsea_upregulated <- gsea[gsea$cluster %in% c(2,6,3),]
terms_upregulated <- create_terms(gsea_upregulated)
circ_upregulated <- circle_dat(terms_upregulated, genes)
```

```{r, fig.height = 10, fig.width = 20, fig.align = "center"}
GOBar(circ_upregulated)
```

Negative Z-score = terms upregulated in ECM branch. Positive Z-score = terms upregulated in metabolic branch.

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_upregulated, labels=2, ID=F)
```

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_upregulated, labels=2, ID=F, table.legend = F)
```

```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
GOBubble(circ_upregulated, labels=8, ID=F, display='mutliple')
```

Relationship between genes and terms

```{r}
#process <- c('extracellular exosome', 'lipid metabolic process', 'purine-containing compound metabolic process', 'fatty acid derivative biosynthetic process')

#genes_ordered <- genes[order(-abs(genes$logFC)),]
#process <- gsea$revigo.description[gsea$cluster == 3]
#chord <- chord_dat(circ_upregulated, genes_ordered)
```


```{r, fig.height = 12, fig.width = 20, fig.align = "center"}
#GOHeat(chord, nlfc = 1)
```

```{r}

```

