---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Notebook for alignment analsyis of the 180504 data.

```{r}
library(Seurat)
library(ggplot2)
```


```{r}
load('output/10x-180504-aligned-metageneplot')
all10x.aligned <- readRDS('output/10x-180504-aligned')
all10x.aligned.ccregout <- readRDS('output/10x-180504-ccregout-aligned')
all10x.aligned.discardedcells <- readRDS('output/10x-180504-cca-discardedcells')
all10x.ccregout.aligned.discardedcells <- readRDS('output/10x-180504-ccregout-cca-discardedcells')
```


Alignment of the data with and without cell cycle effects regressed out. Both were aligned on 30 subspaces, tSNE was performed on the first 15 CCs. 

#Shared correlation per CC

Normal alignment

```{r}
p1
```

Alignment with cell cycle effects regressed out

```{r}
p2
```



#Alignment without cell cycle effects regressed out

tSNE of the aligned data.

```{r}
TSNEPlot(all10x.aligned, group.by='sample_name', pt.size=0.1)
```

tSNE of the aligned data coloured on cell cycle phase.

```{r}
TSNEPlot(all10x.aligned, group.by='Phase', pt.size=0.1)
```

#Alignment with cell cycle regressed out

tSNE of the aligned data with cell cycle effects regressed out.

```{r}
TSNEPlot(all10x.aligned.ccregout, group.by='sample_name', pt.size=0.1)
```

tSNE of the aligned data with cell cycle effects regressed out, colored by phase.

```{r}
TSNEPlot(all10x.aligned.ccregout, group.by='Phase', pt.size=0.1)
```

tSNE of the aligned data with cell cycle effects regressed out, colored by subtissue

```{r}
TSNEPlot(all10x.aligned.ccregout, group.by='depot', pt.size=0.1)
```

#Discarded cells from alignment

Before aligning the samples, sample-specific cells were discarded. Below: discarded cells from normal alignment.

```{r}
discarded <- as.data.frame(table(all10x.aligned.discardedcells@meta.data$sample_name))

names(discarded) <- c('Sample', 'Frequency')

p_discarded <-ggplot(data=discarded, aes(x=Sample, y=Frequency)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle=45, hjust=1)) 

p_discarded
```

Discarded cells from alignment with cell cycle effects regressed out.

```{r}
discarded.ccregout <- as.data.frame(table(all10x.ccregout.aligned.discardedcells@meta.data$sample_name))

names(discarded.ccregout) <- c('Sample', 'Frequency')

p_discarded.ccregout <-ggplot(data=discarded.ccregout, aes(x=Sample, y=Frequency)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle=45, hjust=1)) 

p_discarded.ccregout
```

#Figures for report

```{r echo=T, results='hide', include=F}
fig1 <- plot_grid(
  TSNEPlot(all10x.aligned, group.by='sample_name', pt.size=0.1),
  TSNEPlot(all10x.aligned, group.by='Phase', pt.size=0.1),
  TSNEPlot(all10x.aligned.ccregout, group.by='depot', pt.size=0.1),
  labels=c('a', 'b', 'c'), nrow=3
)
save_plot("plots/180504_alignment.pdf", fig1, base_width=6, base_height=12)
```

```{r fig1, fig.height = 12, fig.width = 6, fig.align = "center"}
fig1
```

Biweight midcorrelation plots.

```{r fig2, fig.height = 4, fig.width = 12, fig.align = "center"}
sfig1 <- plot_grid(
  p1,
  p2,
  labels=c('a', 'b'),
  nrow=1
)
save_plot("plots/supplementary_figures/sfig_180504_biweightplots.pdf", sfig1, base_width=12, base_height=4)
sfig1
```

Discarded cells.

```{r fi3, fig.height = 4, fig.width = 12, fig.align = "center"}
sfig2 <- plot_grid(
  p_discarded,
  p_discarded.ccregout,
  labels=c('a', 'b'),
  nrow=1
)
save_plot("plots/supplementary_figures/sfig_180504_alignment-discardedcells.pdf", sfig2, base_width=12, base_height=4)
sfig2
```

