---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---

This notebook shows the results of different feature selection methods of Monocle. 

```{r results='hide'}
library(monocle)
library(cowplot)
```


#Genes with high dispersion

```{r}
cds_high_disp <- readRDS('output/monocle/180831/monocle_high_dispersion/10x-180831-monocle')
plot_cell_trajectory(cds_high_disp, color_by = 'timepoint')
```

#DE genes between T1T2T3 and T4T5

```{r}
cds_timecombined <- readRDS('output/monocle/180831/monocle_time-combined/10x-180831-monocle')
plot_cell_trajectory(cds_timecombined, color_by='timepoint')
```

#DE genes from cluster resolution 1.5

```{r}
cds_res1.5 <- readRDS('output/monocle/180831/monocle_res1.5/10x-180831-monocle')
plot_cell_trajectory(cds_res1.5, color_by='timepoint')
```

#Dataset split into T1+T2+T3 and T4+T5, DE genes from clusters res0.5.

```{r}
cds_split_res0.5 <- readRDS('output/monocle/180831/monocle_T1T2T3_T4T5_res0.5/10x-180831-noreg-monocle')
plot_cell_trajectory(cds_split_res0.5, color_by='timepoint')
```

#Dataset split into T1+T2+T3 and T4+T5, DE genes from clusters res1.5 are used here.

```{r}
cds <- readRDS('output/monocle/180831/monocle_T1T2T3_T4T5_res1.5/10x-180831-monocle')
plot_cell_trajectory(cds, color_by='timepoint')
```

With doublets removed (used for all further analyses).

```{r}
cds <- readRDS('output/monocle/180831/10x-180831-monocle-monocle_genelist_T1T2T3_T4T5_res.1.5')
plot_cell_trajectory(cds, color_by='timepoint')
```


#DE genes between brown and white.

```{r}
cds_type <- readRDS('output/monocle/180831/monocle_type/10x-180831-monocle')
plot_cell_trajectory(cds_type, color_by='timepoint')
```

#DE genes between top10% lowest PC2 score and top10% highest PC2 score.

```{r fig17, fig.height = 10, fig.width = 12, fig.align = "center"}
cds_pc2 <- readRDS('output/monocle/180831/monocle_PC2/10x-180831-monocle')
plot_grid(
  plot_cell_trajectory(cds_pc2, color_by='timepoint'),
  plot_cell_trajectory(cds_pc2, color_by='State.old') + scale_color_manual(values=c("#f67770", "#964B00", "orange"), name = "State"),
  plot_cell_trajectory(cds_pc2, color_by='Pseudotime'),
  plot_cell_trajectory(cds_pc2, color_by='State'),
  ncol=2
)
```
