---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---

Analysis of the 10x-180504 samples.

Loading the required packages and datasets. 

```{r}
library(Seurat)
library(ggplot2)
library(dplyr)

all10x <- readRDS('output/10x-180504')
all10x.ccregout <- readRDS('output/10x-180504-ccregout')
```

#QC Plots

```{r}
VlnPlot(all10x, features.plot='nGene', group.by='sample_name', point.size.use=-1, x.lab.rot=T)
```

```{r}
VlnPlot(all10x, features.plot='nUMI', group.by='sample_name', point.size.use=-1, x.lab.rot=T)
```

```{r}
VlnPlot(all10x, features.plot='percent.mito', group.by='sample_name', point.size.use=-1, x.lab.rot=T)
```

```{r}
GenePlot(all10x, 'nUMI', 'nGene')
```


#TSNE 

Below are several tSNE plots of the 10x-180504 data. tSNE was performed on the first 15 principal components of the log-normalized scaled (nUMI and percent.mito regressed out) data.

Visceral and perirenal seem a bit mixed, and supraclavicular and subcutaneous too. 

```{r}
TSNEPlot(all10x, pt.size=0.1, group.by='sample_name', do.label=T)
```

tSNE plots of samples within their depot.
Peri2 and Peri3 seem to overlap really well, as well as Supra1 and Supra2, and Visce1 and Visce3.

```{r echo=T, results='hide', include=FALSE}
subtissues <- all10x@meta.data$depot
t1 <- TSNEPlot(all10x, cells.use=all10x@cell.names[which(subtissues == 'Peri')], group.by='sample_name', pt.size=0.2)
t2 <- TSNEPlot(all10x, cells.use=all10x@cell.names[which(subtissues == 'Subq')], group.by='sample_name', pt.size=0.2)
t3 <- TSNEPlot(all10x, cells.use=all10x@cell.names[which(subtissues == 'Supra')], group.by='sample_name', pt.size=0.2)
t4 <- TSNEPlot(all10x, cells.use=all10x@cell.names[which(subtissues == 'Visce')], group.by='sample_name', pt.size=0.2)
```


```{r}
plot_grid(t1, t2, t3, t4)
```

tSNE colored on subtissue.

```{r}
TSNEPlot(all10x, group.by='depot', pt.size=0.1)
```

tSNE colored by cell cycle phase. 

```{r}
TSNEPlot(all10x, group.by='Phase', pt.size=0.1)
```

#Clusters 

Some clustering with different resolutions.
res=0.5

```{r}
TSNEPlot(all10x, pt.size=0.1, group.by='res.0.5', do.label=T)
```

res=0.7

```{r}
TSNEPlot(all10x, pt.size=0.1, group.by='res.0.7', do.label=T)
```

res=1

```{r}
TSNEPlot(all10x, pt.size=0.1, group.by='res.1', do.label=T)
```


#Cell cycle regression

T-SNE of the data with cell cycle effects regressed out. There does not seem to be a lot of structure within clusters now. 

```{r}
TSNEPlot(all10x.ccregout, pt.size=0.1, group.by='sample_name')
```

No cell cycle effect anymore. 

```{r}
TSNEPlot(all10x.ccregout, pt.size=0.1, group.by='Phase')
```

Subtissues

```{r echo=T, results='hide', include=FALSE}
subtissues <- all10x.ccregout@meta.data$depot
t1 <- TSNEPlot(all10x.ccregout, cells.use=all10x.ccregout@cell.names[which(subtissues %in% 'Peri')], group.by='sample_name', pt.size=0.2)
t2 <- TSNEPlot(all10x.ccregout, cells.use=all10x.ccregout@cell.names[which(subtissues %in% 'Subq')], group.by='sample_name', pt.size=0.2)
t3 <- TSNEPlot(all10x.ccregout, cells.use=all10x.ccregout@cell.names[which(subtissues %in% 'Supra')], group.by='sample_name', pt.size=0.2)
t4 <- TSNEPlot(all10x.ccregout, cells.use=all10x.ccregout@cell.names[which(subtissues %in% 'Visce')], group.by='sample_name', pt.size=0.2)

```

```{r}
plot_grid(t1, t2, t3, t4)
```

```{r}
TSNEPlot(all10x.ccregout, pt.size=0.1, group.by='depot')
```

#PCA

Some PCA plots. PC1 seems to capture cell cycle effects, and PC2 seems to capture some of the sample variability. 

```{r}
PCAPlot(all10x, group.by='Phase', pt.size=0.1)
```

```{r}
PCAPlot(all10x, group.by='sample_name', pt.size=0.1)
```

PCA plot of the cell cycle regressed out data. There is no cell cycle effect anymore. 

```{r}
PCAPlot(all10x.ccregout, group.by='Phase', pt.size=0.1)
```

```{r}
PCAPlot(all10x.ccregout, group.by='sample_name', pt.size=0.1)
```


#Metadata plots


```{r}
FeaturePlot(all10x, c("nGene"), cols.use = c("grey","blue"), no.legend=F)
```

```{r}
FeaturePlot(all10x, c("percent.mito"), cols.use = c("grey","blue"), no.legend=F)
```


```{r}
FeaturePlot(all10x, c("nUMI"), cols.use = c("grey","blue"), no.legend=F)
```

Diff

```{r}
TSNEPlot(all10x, group.by='diff', pt.size=0.1)
```

```{r}
all10x@meta.data['diff_int'] <- unlist(lapply(as.vector(unlist(all10x@meta.data$diff)), function(x){return(strtoi(strsplit(x, '%')))}))
FeaturePlot(all10x, features.plot='diff_int', cols.use=c('gray', 'blue'), no.legend=F)
```

ucp1.ctrl

```{r}
TSNEPlot(all10x, group.by='ucp1.ctrl', pt.size=0.1)
```

ucp1.ne

```{r}
TSNEPlot(all10x, group.by='ucp1.ne', pt.size=0.1)
```

bmi

```{r}
TSNEPlot(all10x, group.by='bmi', pt.size=0.1)
```

age 

```{r}
TSNEPlot(all10x, group.by='age', pt.size=0.1)
```


```{r}
VlnPlot(all10x, group.by='sample_name', features.plot=c('nGene'), point.size.use = -1, x.lab.rot=T)
```

```{r}
VlnPlot(all10x, group.by='sample_name', features.plot=c('nUMI'), point.size.use = -1, x.lab.rot=T)
```

```{r}
VlnPlot(all10x, group.by='sample_name', features.plot=c('percent.mito'), point.size.use = -1, x.lab.rot=T)
```

#Mixture cluster 12

Sample composition in cluster 12. 

```{r}
cluster12 <- SubsetData(all10x, cells.use=rownames(all10x@meta.data)[which(all10x@meta.data$res.0.5 %in% 12)])
rotate_x <- function(data, column_to_plot, labels_vec, rot_angle) {
     plt <- barplot(data[[column_to_plot]], col='steelblue', xaxt="n")
     text(plt, par("usr")[3], labels = labels_vec, srt = rot_angle, adj = c(1.1,1.1), xpd = TRUE, cex=1)
}
rotate_x((cluster12@meta.data %>% count(sample_name))[,2], 'n', as.vector(unlist((cluster12@meta.data %>% count(sample_name))[,1])), 45)
```

#Figures for report

```{r echo=T, results='hide', include=F}
fig1 <- plot_grid(
  PCAPlot(all10x, group.by='Phase', pt.size=0.1),
  TSNEPlot(all10x, group.by='sample_name', pt.size=0.1),
  TSNEPlot(all10x, group.by='Phase', pt.size=0.1),
  TSNEPlot(all10x.ccregout, group.by='sample_name', pt.size=0.1),
  labels = "auto", nrow = 2)
save_plot("plots/180504_pca_tsne.pdf", fig1, base_width=12, base_height = 9)
```


```{r fig1, fig.height = 9, fig.width = 12, fig.align = "center"}
fig1
```

```{r fig2, fig.height = 3, fig.width = 12, fig.align = "center"}
#Supplementary figures
sfig1 <- plot_grid(
  VlnPlot(all10x, group.by='sample_name', features.plot=c('nGene'), point.size.use = -1, x.lab.rot=T, size.x.use=8),
  VlnPlot(all10x, group.by='sample_name', features.plot=c('nUMI'), point.size.use = -1, x.lab.rot=T, size.x.use=8),
  VlnPlot(all10x, group.by='sample_name', features.plot=c('percent.mito'), point.size.use = -1, x.lab.rot=T, size.x.use=8), 
  labels="auto", nrow=1
)
save_plot("plots/supplementary_figures/sfig_180504_qcplots.pdf", sfig1, base_width=12, base_height=3)
sfig1
```

```{r fig3, fig.height = 3, fig.width = 12, fig.align = "center"}
sfig2 <- plot_grid(
  PCElbowPlot(all10x, num.pc=30),
  PCElbowPlot(all10x.ccregout, num.pc=30),
  labels="auto", nrow=1)
save_plot("plots/supplementary_figures/sfig_180504_pcelbow.pdf", sfig2, base_width=12, base_height=5)
sfig2
```

