---
title: "R Notebook"
output: html_notebook
---

```{r message=F}
library(Seurat)
library(dplyr)
library(tibble)
library(cowplot)
library(DT)
```

Data loading and processing

```{r}
# file.data <- "/nfsdata/projects/timshel/sc-scheele_lab_adipose_fluidigm_c1/data-vijay/GSE129363_SVF_Normalised_Data.txt.gz"
# df.data <- read_tsv(file.data) %>% rename(gene=X1)# WORKS and parsed correctly. Warning message: Missing column names filled in: 'X1' [1] 
# df.data <- data.table::fread(file.data, nThread=24, showProgress=T)
# df.data <- df.data %>% column_to_rownames(var="V1")
# 
# ### Meta-data
# file.metadata <- "/nfsdata/projects/timshel/sc-scheele_lab_adipose_fluidigm_c1/data-vijay/GSE129363_cell_metadata.txt.gz"
# df.metadata <- read_tsv(file.metadata)
# df.metadata <- df.metadata %>% rename(cell_barcode=X1, diabetes=Condition, depot=Tissue, celltype=ClsID)
# # *TODO*: rename IA-->visc; SC-->subq
# df.metadata <- df.metadata %>% column_to_rownames(var="cell_barcode")
# 
# ### Create object
# vijay <- CreateSeuratObject(counts=df.data, meta.data=df.metadata, project="vijay") # Warning: Feature names cannot have underscores ('_'), replacing with dashes ('-')
# vijay[["RNA"]]@data <- vijay[["RNA"]]@counts # copy slot because data is already normalized
# # NB: do not perform further (log)normalization as this has already been done
# 
# vijay <- ScaleData(vijay)
# vijay <- FindVariableFeatures(vijay)
# vijay <- RunPCA(vijay)
# #ElbowPlot(vijay, reduction='pca')
# vijay <- RunTSNE(vijay, dims=1:11)
# vijay <- RunUMAP(vijay, dims=1:11)
# saveRDS(vijay, 'output/seurat_objects/vijay/vijay.rds')
```

```{r}
vijay <- readRDS('output/seurat_objects/vijay/vijay.rds')
```

#Metadata plots

```{r fig.height = 10, fig.width = 12, fig.align = "center"}
#plot cell types + overlayed with meta data
plot_grid(
 UMAPPlot(vijay, group.by='celltype', label=T), 
 UMAPPlot(vijay, group.by='depot'),
 UMAPPlot(vijay, group.by='diabetes')
)
```

```{r fig.height = 5, fig.width = 12, fig.align = "center"}
#plot cell types + overlayed with meta data
plot_grid(
 FeaturePlot(vijay, features='nFeature_RNA'), 
 FeaturePlot(vijay, features='nCount_RNA')
)
```

#U/L branch marker genes

Plot the U and L branch marker genes. 

```{r}
data_180831 <- readRDS('output/seurat_objects/180831/10x-180831-S3')
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_u_l <- read.table('output/markergenes/180831/markers_10x-180831_upperbranch_lowerbranch_negbinom', sep='\t', header=T)
markers_u <- markers_u_l[order(-markers_u_l$avg_logFC),]
markers_l <- markers_u_l[order(markers_u_l$avg_logFC),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

And how are they expressed in the Vijay data?

```{r}
UMAPPlot(vijay, group.by='celltype', label=T)
```


```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(vijay, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

Some U branch markers are not really expressed (SCD, PLIN4, GDP1, RBP4). Interestingly most U branch markers are also expressed in I1 and I3 and E1-E3. 

#Seurat integration with 10x-180831 data

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
#anchors <- FindIntegrationAnchors(object.list = list(vijay, data_180831), dims = 1:20)
#integrated <- IntegrateData(anchorset = anchors, dims = 1:20)
#integrated <- ScaleData(integrated, verbose = FALSE)
#integrated <- RunPCA(integrated, npcs = 30, verbose = FALSE)
#integrated <- RunUMAP(integrated, reduction = "pca", dims = 1:10)
#integrated <- RunTSNE(integrated, reduction = "pca", dims = 1:10)
#saveRDS(integrated, '/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/vijay/vijay.180831.integrated.rds')

integrated <- readRDS('output/seurat_objects/vijay/vijay.180831.integrated.rds')
integrated@meta.data['dataset'] <- '10x-180831'
integrated@meta.data[which(is.na(integrated@meta.data$branch)), 'dataset'] <- 'Vijay'

plot_grid(
  UMAPPlot(integrated, group.by='dataset'),
  UMAPPlot(integrated, group.by='celltype', label=T),
  UMAPPlot(integrated, group.by='branch'), ncol=2
)
```

Structural cells overlap with clusters P4, P5, P6, P2, P7. Metabolic cells overlap partly with P2 and P7 and also I3.  

#Predict cell types from 10x-180831 data

```{r}
#find anchors
anchors <- FindTransferAnchors(reference = data_180831, query = vijay, dims = 1:20)
anchors_cca <- FindTransferAnchors(reference = data_180831, query = vijay, dims = 1:20, reduction = 'cca')

#transfer labels
predictions_pca_project <- TransferData(anchors, data_180831$branch, dims = 1:20, weight.reduction='pcaproject')
predictions_pca <- TransferData(anchors, data_180831$branch, dims=1:20, weight.reduction='pca')
predictions_cca <- TransferData(anchors_cca, data_180831$branch, dims=1:20, weight.reduction='cca')

#rename colnames
names(predictions_pca_project) <- unlist(lapply(names(predictions_pca_project), function(x){return(paste('pca_project.', x, sep=''))}))
names(predictions_pca) <- unlist(lapply(names(predictions_pca), function(x){return(paste('pca.', x, sep=''))}))
names(predictions_cca) <- unlist(lapply(names(predictions_cca), function(x){return(paste('cca.', x, sep=''))}))

vijay <- AddMetaData(vijay, metadata = predictions_pca_project)
vijay <- AddMetaData(vijay, metadata = predictions_pca)
vijay <- AddMetaData(vijay, metadata = predictions_cca)
```

Scores

```{r, fig.height = 30, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(vijay, group.by='celltype', label=T),
  FeaturePlot(vijay, features='pca.prediction.score.Progenitor'),
  FeaturePlot(vijay, features='pca_project.prediction.score.Progenitor'),
  FeaturePlot(vijay, features='cca.prediction.score.Progenitor'),
  FeaturePlot(vijay, features='pca.prediction.score.Metabolic'),
  FeaturePlot(vijay, features='pca_project.prediction.score.Metabolic'),
  FeaturePlot(vijay, features='cca.prediction.score.Metabolic'),
  FeaturePlot(vijay, features='pca.prediction.score.ECM'),
  FeaturePlot(vijay, features='pca_project.prediction.score.ECM'),
  FeaturePlot(vijay, features='cca.prediction.score.ECM'),
  UMAPPlot(vijay, group.by='pca.predicted.id'),
  UMAPPlot(vijay, group.by='pca_project.predicted.id'),
  UMAPPlot(vijay, group.by='cca.predicted.id'),
  ncol=2
)
```

Assign the labels if threshold is above 0.7 (same threshold as the wolfrum label transfer). 

```{r}
assign_labels <- function(colname, threshold=0.7){
  pred_ids <- unlist(as.vector(apply(vijay@meta.data[,c(paste(colname,'.prediction.score.max', sep=''), paste(colname, '.predicted.id', sep=''))], 1, function(x){
    if (x[[1]] < threshold){
      return(NA)
    } else{
      return(x[[2]])
    }
  })))
  return(pred_ids)
}

for (col in c('pca_project', 'pca', 'cca')){
  for (t in c(0.5, 0.7, 0.9, 0.95, 0.99)){
    preds <- assign_labels(col, t)
    vijay <- AddMetaData(vijay, preds, col.name=paste(col, 'predicted_label', t, sep='.'))
  }
}
```

Threshold for prediction = 0.5

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(vijay, group.by='pca.predicted_label.0.5'),
  UMAPPlot(vijay, group.by='pca_project.predicted_label.0.5'),
  UMAPPlot(vijay, group.by='cca.predicted_label.0.5'), ncol=2
)
```

Threshold for prediction = 0.7

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(vijay, group.by='pca.predicted_label.0.7'),
  UMAPPlot(vijay, group.by='pca_project.predicted_label.0.7'),
  UMAPPlot(vijay, group.by='cca.predicted_label.0.7'), ncol=2
)
```

Threshold for prediction = 0.9

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(vijay, group.by='pca.predicted_label.0.9'),
  UMAPPlot(vijay, group.by='pca_project.predicted_label.0.9'),
  UMAPPlot(vijay, group.by='cca.predicted_label.0.9'), ncol=2
)
```

I3 = adipose tissue macrophages (Vijay paper). Perhaps no metabolic preadipocytes because of maturation?

```{r}
UMAPPlot(vijay, group.by='celltype', label=T)
```


