---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---


```{r message=F}
library(liger)
library(Seurat)
library(Seurat)
library(SeuratWrappers)
library(monocle)
library(cowplot)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(DT)
library(harmony)
```

```{r}
wolfrum <- readRDS('/projects/timshel/sc-scheele_lab_adipose_fluidigm_c1/data-wolfrum/wolfrum.compute.seurat_obj.rds')
data_180831 <- readRDS('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/180831/10x-180831-S3')
```

# Which clusters are adipocytes/preadipocytes?

```{r fig.height = 4, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum, group.by='orig.ident', label=T),
  UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
)
```

```{r}
markers <- read.table('output/markergenes/wolfrum/markers_wolfrum.compute.seurat_obj.rds_seurat_clusters_negbinom', sep='\t', header=T)
```

Top 10 positive markers per cluster 

```{r}
pos_markers_top10 <- markers %>% 
  group_by(cluster) %>% 
  top_n(n=10, wt=avg_logFC)
  
neg_markers_top20 <- markers %>% 
  group_by(cluster) %>% 
  top_n(n=6, wt=desc(avg_logFC))

pos_markers_top10
```

Which clusters are the preadipocytes? Check by plotting some of the DE genes between T1T2T3 and T4T5 in the 180831 dataset.

```{r}
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_T1T2T3_T4T5 <- read.table('output/markergenes/180831/markers_10x-180831_time_combined_negbinom', header=T)
markers_T1T2T3_T4T5 <- markers_T1T2T3_T4T5[order(-markers_T1T2T3_T4T5$avg_logFC),]
markers_T1T2T3 <- markers_T1T2T3_T4T5[which(markers_T1T2T3_T4T5$cluster == 1),]
markers_T4T5 <- markers_T1T2T3_T4T5[which(markers_T1T2T3_T4T5$cluster == 2),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_T1T2T3$gene)[1:10], as.vector(markers_T4T5$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

How are they expressed in the Wolfrum data?

```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(wolfrum, features=c(as.vector(markers_T1T2T3$gene)[1:10], as.vector(markers_T4T5$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

Not all genes are expressed clearly in specific clusters. It looks like cluster 22, 21, 5, 14, 23, 11 and 10 are the preadipocytes. 

```{r}
UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
```

Also plot the top 10 ECM and Metabolic markers.

```{r}
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_u_l <- read.table('output/markergenes/180831/markers_10x-180831_upperbranch_lowerbranch_negbinom', sep='\t', header=T)
markers_u <- markers_u_l[order(-markers_u_l$avg_logFC),]
markers_l <- markers_u_l[order(markers_u_l$avg_logFC),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

And how are they expressed in the Wolfrum data?

```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(wolfrum, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

The U and L branch markers are clearly expressed in the two clusters in the top right of the UMAP plot. 

Check how many of the marker genes are found in each cluster. 

```{r}
get_gene_overlap_per_cluster <- function(min_logFC, n_genes){
  pos_markers <- markers[markers$avg_logFC > min_logFC,]

  genes_P <- c(as.vector(markers_T1T2T3$gene)[1:n_genes])
  #genes_P <- c(as.vector(markers_T1T2T3$gene)[1:n_genes], as.vector(markers_T4T5$gene)[1:n_genes])
  genes_U <- as.vector(markers_u$gene)[1:n_genes]
  genes_L <- as.vector(markers_l$gene)[1:n_genes]
  all_genes <- unique(c(genes_P, genes_U, genes_L))
  
  df <- as.data.frame(matrix(ncol=4, nrow=length(unique(pos_markers$cluster))))
  colnames(df) <- c('total_overlap', 'overlap_P', 'overlap_L', 'overlap_U')
  rownames(df) <- unique(pos_markers$cluster)

  for (i in 1:length(unique(pos_markers$cluster))){
    genes_cluster <- as.vector(pos_markers[pos_markers$cluster == i,]$gene)
    df[i, 'overlap_P'] <- round(length(intersect(genes_cluster, genes_P)) / length(genes_P), 2)
    df[i, 'overlap_L'] <- round(length(intersect(genes_cluster, genes_L)) / length(genes_L), 2)
    df[i, 'overlap_U'] <- round(length(intersect(genes_cluster, genes_U)) / length(genes_U), 2)
    df[i, 'total_overlap'] <- round(length(intersect(genes_cluster, all_genes)) / length(all_genes), 2)
  }
  return(df)
}

print_top_clusters <- function(df){
  print(paste('P: ', toString(rownames(df[order(-df$overlap_P),][1:3,])), sep=''))
  print(paste('L: ', toString(rownames(df[order(-df$overlap_L),][1:3,])), sep=''))
  print(paste('U: ', toString(rownames(df[order(-df$overlap_U),][1:3,])), sep=''))
}

```

Table shows the percentage of genes with avgLogFC > 0.7 that was found in the cluster. Sort on columns to get the top clusters.

```{r}
logfc_0.5_genes_100 <- get_gene_overlap_per_cluster(0.5, 100)
datatable(logfc_0.5_genes_100)
```

Check the top 5 clusters per branch for different n_genes and min logFC

```{r}
logfc <- c(0.25, 0.5, 0.7)
n_genes <- c(100, 100, 100)
for (i in 1:length(logfc)){
  print(paste('min_logFC: ', logfc[i], ' | n_genes: ', n_genes[i], sep=''))
  df <- get_gene_overlap_per_cluster(logfc[i], n_genes[i])
  print_top_clusters(df)
}
```

There is some overlap between branches which is good. Cluster 11 is shared between P and L, cluster 14 is shared between P and U and cluster 23 is shared between L and U. This would indicate that cluster 24 contains most immature preadipocytes, cluster 10 contains most mature L branch cells and cluster 22 contains most mature U branch cells.\\

Based on the results:\\
P = 24\\
L = 11\\
U = 14\\

Hypothesis: cluster 23 represents preadipocytes at the start of differentation (the cell states between T3 and T4 in 180831 data that we missed). Cluster 5 represents even more mature metabolic cells and cluster 11 represents more mature ECM cells.\\

Clusters 22 shares most genes with the P branch. Cluster 23 most with the U branch. (see datatable above). These could also represent the preadipocytes at start of differentiation or the cells that transfer back to progenitor cells. \\

```{r}
UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
```

#Preadipocyte subset

```{r}
#Idents(wolfrum) <- wolfrum@meta.data$seurat_clusters
#preadipocyte_subset <- subset(wolfrum, idents=c(5, 14, 23, 11, 10, 21, 22, 24))
#preadipocyte_subset <- FindVariableFeatures(preadipocyte_subset)
#preadipocyte_subset <- ScaleData(preadipocyte_subset)
#preadipocyte_subset <- RunPCA(object=preadipocyte_subset, npcs=30)
#ElbowPlot(preadipocyte_subset, ndims=30)
#preadipocyte_subset <- FindNeighbors(object = preadipocyte_subset, dims=1:13)
#preadipocyte_subset <- FindClusters(object = preadipocyte_subset, resolution=0.8)
#preadipocyte_subset <- RunTSNE(object = preadipocyte_subset, dims=1:13)
#preadipocyte_subset <- RunUMAP(object = preadipocyte_subset, dims=1:13)
#saveRDS(preadipocyte_subset, '/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.rds')
preadipocyte_subset <- readRDS('output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.rds')

plot_grid(
  UMAPPlot(preadipocyte_subset, group.by='all_data_seurat_clusters', label=T),
  TSNEPlot(preadipocyte_subset, group.by='all_data_seurat_clusters', label=T)
)
```


# Figures

```{r}
adipoq <- FeaturePlot(wolfrum, features='ADIPOQ') + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

lipe <- FeaturePlot(wolfrum, features='LIPE') + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

apod <- FeaturePlot(wolfrum, features='APOD') + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

dcn <- FeaturePlot(wolfrum, features='DCN') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5)) 
```

```{r, fig.height = 4, fig.width = 16, fig.align = "center"}
g <- plot_grid(
  adipoq, lipe, apod, dcn, ncol=4,  rel_widths=c(1, 1, 1, 1.3)
)
g
```

```{r}
save_plot("figures/figures_paper/main_figures/Figure_wolfrum/featureplots.pdf", g, base_width=16, base_height=4)
```

EBF2 and LEP

```{r, fig.height = 4, fig.width = 12, fig.align = "center"}
ebf2 <- FeaturePlot(wolfrum, features='EBF2') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

pparg <- FeaturePlot(wolfrum, features='PPARG') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

lep <- FeaturePlot(wolfrum, features='LEP') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

ucp1 <- FeaturePlot(wolfrum, features='UCP1') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

ebf2vln <- VlnPlot(wolfrum, features='EBF2', group.by='RNA_snn_res.0.8', pt.size=-1) + NoLegend()

lepvln <- VlnPlot(wolfrum, features='LEP', group.by='RNA_snn_res.0.8', pt.size=0.1) + NoLegend()

ucp1vln <- VlnPlot(wolfrum, features='UCP1', group.by='RNA_snn_res.0.8', pt.size=0.1) + NoLegend()

ppargvln <- VlnPlot(wolfrum, features='PPARG', group.by='RNA_snn_res.0.8', pt.size=0.1) + NoLegend()

plot_grid(ebf2vln, lepvln, ucp1vln, ppargvln, ncol=2)
```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ebf2.pdf", ebf2, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ebf2vln.pdf", ebf2vln, base_width=6, base_height=2)

#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/pparg.pdf", pparg, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ppargvln.pdf", ppargvln, base_width=6, base_height=2)

#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ucp1.pdf", ucp1, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ucp1vln.pdf", ucp1vln, base_width=6, base_height=2)

#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/lep.pdf", lep, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/lepvln.pdf", lepvln, base_width=5, base_height=2)

```


```{r}
UMAPPlot(wolfrum, label=T, group.by='RNA_snn_res.0.8') 
```

```{r, fig.height = 6, fig.width = 12, fig.align = "center"}
VlnPlot(wolfrum, features=c('EBF2', 'PPARG'), group.by='RNA_snn_res.0.8', pt.size=-1, ncol=1)
```


```{r, fig.height = 6, fig.width = 6, fig.align = "center"}
#EBF2 and PPARG
#PPARG: 5, 14, 23
#EBF2: 10, 11, 21, 23

new_labels <- unlist(lapply(wolfrum@meta.data$RNA_snn_res.0.8, function(x){
  if (x %in% c(5, 14, 23, 10, 11, 21, 23)){
    return(x)
  } else {
    return(NA)
  }
}))

wolfrum@meta.data['labels_clusters_preadipocytes'] <- new_labels
p <- UMAPPlot(wolfrum, group.by='labels_clusters_preadipocytes', label=T) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm")) + NoLegend()
p
```

```{r}
save_plot("figures/figures_paper/main_figures/Figure_wolfrum/UMAP_adipocyte_clusters2.pdf", p, base_width=6, base_height=5)
```


```{r, fig.height = 4, fig.width = 8.5, fig.align = "center"}
ebf2 <- FeaturePlot(wolfrum, features='EBF2', pt.size=0.5) + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

pparg <- FeaturePlot(wolfrum, features='PPARG', pt.size=0.5) + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=15), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5)) 

g <- plot_grid(
  ebf2, pparg, ncol=2, rel_widths = c(1, 1.3)
)

g
```

```{r}
save_plot("figures/figures_paper/main_figures/Figure_wolfrum/EBF2_PPARG_ptsize0.5.pdf", g, base_width=8.5, base_height=4)
```

```{r, fig.height = 4, fig.width = 8.5, fig.align = "center"}
pparg <- FeaturePlot(data_180831, features='PPARG', pt.size=0.5) + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,2.5))

ebf2 <- FeaturePlot(data_180831, features='EBF2', pt.size=0.5) + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=15), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,2.5)) 

g <- plot_grid(
  pparg, ebf2, ncol=2, rel_widths = c(1, 1.4)
)

g
```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/EBF2_PPARG_no-axes.pdf", g, base_width=8.5, base_height=4)
```

```{r}
data.180831.brown <- subset(data_180831, cells=rownames(data_180831@meta.data)[data_180831@meta.data$type == 'brown'])
```


```{r, fig.height = 3, fig.width = 8.5, fig.align = "center"}
colors.states.labels <- c(
  P='#ecdd83',
  L='#93c8bc',
  U='#e27268')

ebf2_vln_brown <- VlnPlot(data.180831.brown, features='EBF2', group.by='State.labels', pt.size=-1, cols=colors.states.labels) + NoLegend() + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 0, size=15), plot.title = element_text(size=20), axis.text.y = element_text(size=15), plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines"))  + scale_y_continuous(breaks=c(0, 1, 2))

pparg_vln_brown <- VlnPlot(data.180831.brown, features='PPARG', group.by='State.labels', pt.size=-1, cols=colors.states.labels) + NoLegend() + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 0, size=15), plot.title = element_text(size=20), axis.text.y = element_text(size=15), plot.margin = unit(c(1.5,1.5,1.5,1.5), "lines")) + scale_y_continuous(breaks=c(0, 1, 2))

g <- plot_grid(
  pparg_vln_brown, ebf2_vln_brown, ncol=2
)

g

```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/EBF2_PPARG_brown_violinplots.pdf", g, base_width=8.5, base_height=3)
```

```{r, fig.height = 10, fig.width = 8, fig.align = "center"}
plots <- VlnPlot(wolfrum, features=c('ADIPOQ', 'LIPE', 'PLIN4', 'FABP4', 'ADIRF', 'APOD', 'MGP', 'DCN', 'CCDC80', 'PLAC9'), group.by='RNA_snn_res.0.8', pt.size=-1, combine=F)

for (i in 1:length(plots)){
  if (i == length(plots)){
    plots[[i]] <- plots[[i]] + NoLegend() + 
      theme(plot.title=element_blank(), 
            axis.title.y=element_blank(), 
            axis.line.x=element_blank(),
            axis.text.x=element_text(angle=0, size=12),
            plot.margin = unit(c(0, 0, 0, 0), "cm")) + 
      labs(x='Cluster')
  } else {
    plots[[i]] <- plots[[i]] + NoLegend() + 
      theme(plot.title=element_blank(), 
            axis.title.y=element_blank(), 
            axis.line.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_blank(),
            axis.title.x=element_blank(),
            plot.margin = unit(c(0, 0, 0, 0), "cm"))
  }
}

vlnplts <- plot_grid(plotlist=plots, ncol=1, rel_heights=c(1,1,1,1,1,1,1,1,1,1.6))
vlnplts
```

