---
title: "R Notebook"
output: html_notebook
---


```{r message=F}
library(Seurat)
library(monocle)
library(cowplot)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(DT)
```

```{r}
wolfrum <- readRDS('/projects/timshel/sc-scheele_lab_adipose_fluidigm_c1/data-wolfrum/wolfrum.compute.seurat_obj.rds')
data_180831 <- readRDS('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/180831/10x-180831-S3')
```



#Which clusters are adipocytes/preadipocytes?

```{r fig.height = 4, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum, group.by='orig.ident', label=T),
  UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
)
```

```{r}
markers <- read.table('output/markergenes/wolfrum/markers_wolfrum.compute.seurat_obj.rds_seurat_clusters_negbinom', sep='\t', header=T)
```

Top 10 positive markers per cluster 

```{r}
pos_markers_top10 <- markers %>% 
  group_by(cluster) %>% 
  top_n(n=10, wt=avg_logFC)
  
neg_markers_top20 <- markers %>% 
  group_by(cluster) %>% 
  top_n(n=6, wt=desc(avg_logFC))

pos_markers_top10
```

Which clusters are the preadipocytes? Check by plotting some of the DE genes between T1T2T3 and T4T5 in the 180831 dataset.

```{r}
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_T1T2T3_T4T5 <- read.table('output/markergenes/180831/markers_10x-180831_time_combined_negbinom', header=T)
markers_T1T2T3_T4T5 <- markers_T1T2T3_T4T5[order(-markers_T1T2T3_T4T5$avg_logFC),]
markers_T1T2T3 <- markers_T1T2T3_T4T5[which(markers_T1T2T3_T4T5$cluster == 1),]
markers_T4T5 <- markers_T1T2T3_T4T5[which(markers_T1T2T3_T4T5$cluster == 2),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_T1T2T3$gene)[1:10], as.vector(markers_T4T5$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

How are they expressed in the Wolfrum data?

```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(wolfrum, features=c(as.vector(markers_T1T2T3$gene)[1:10], as.vector(markers_T4T5$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

Not all genes are expressed clearly in specific clusters. It looks like cluster 22, 21, 5, 14, 23, 11 and 10 are the preadipocytes. 

```{r}
UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
```

Also plot the top 10 ECM and Metabolic markers.

```{r}
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_u_l <- read.table('output/markergenes/180831/markers_10x-180831_upperbranch_lowerbranch_negbinom', sep='\t', header=T)
markers_u <- markers_u_l[order(-markers_u_l$avg_logFC),]
markers_l <- markers_u_l[order(markers_u_l$avg_logFC),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

And how are they expressed in the Wolfrum data?

```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(wolfrum, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

The U and L branch markers are clearly expressed in the two clusters in the top right of the UMAP plot. 

Check how many of the marker genes are found in each cluster. 

```{r}
get_gene_overlap_per_cluster <- function(min_logFC, n_genes){
  pos_markers <- markers[markers$avg_logFC > min_logFC,]

  genes_P <- c(as.vector(markers_T1T2T3$gene)[1:n_genes])
  #genes_P <- c(as.vector(markers_T1T2T3$gene)[1:n_genes], as.vector(markers_T4T5$gene)[1:n_genes])
  genes_U <- as.vector(markers_u$gene)[1:n_genes]
  genes_L <- as.vector(markers_l$gene)[1:n_genes]
  all_genes <- unique(c(genes_P, genes_U, genes_L))
  
  df <- as.data.frame(matrix(ncol=4, nrow=length(unique(pos_markers$cluster))))
  colnames(df) <- c('total_overlap', 'overlap_P', 'overlap_L', 'overlap_U')
  rownames(df) <- unique(pos_markers$cluster)

  for (i in 1:length(unique(pos_markers$cluster))){
    genes_cluster <- as.vector(pos_markers[pos_markers$cluster == i,]$gene)
    df[i, 'overlap_P'] <- round(length(intersect(genes_cluster, genes_P)) / length(genes_P), 2)
    df[i, 'overlap_L'] <- round(length(intersect(genes_cluster, genes_L)) / length(genes_L), 2)
    df[i, 'overlap_U'] <- round(length(intersect(genes_cluster, genes_U)) / length(genes_U), 2)
    df[i, 'total_overlap'] <- round(length(intersect(genes_cluster, all_genes)) / length(all_genes), 2)
  }
  return(df)
}

print_top_clusters <- function(df){
  print(paste('P: ', toString(rownames(df[order(-df$overlap_P),][1:3,])), sep=''))
  print(paste('L: ', toString(rownames(df[order(-df$overlap_L),][1:3,])), sep=''))
  print(paste('U: ', toString(rownames(df[order(-df$overlap_U),][1:3,])), sep=''))
}

```

Table shows the percentage of genes with avgLogFC > 0.7 that was found in the cluster. Sort on columns to get the top clusters.

```{r}
logfc_0.5_genes_100 <- get_gene_overlap_per_cluster(0.5, 100)
datatable(logfc_0.5_genes_100)
```

Check the top 5 clusters per branch for different n_genes and min logFC

```{r}
logfc <- c(0.25, 0.5, 0.7)
n_genes <- c(100, 100, 100)
for (i in 1:length(logfc)){
  print(paste('min_logFC: ', logfc[i], ' | n_genes: ', n_genes[i], sep=''))
  df <- get_gene_overlap_per_cluster(logfc[i], n_genes[i])
  print_top_clusters(df)
}
```

There is some overlap between branches which is good. Cluster 11 is shared between P and L, cluster 14 is shared between P and U and cluster 23 is shared between L and U. This would indicate that cluster 24 contains most immature preadipocytes, cluster 10 contains most mature L branch cells and cluster 22 contains most mature U branch cells.\\

Based on the results:\\
P = 24\\
L = 11\\
U = 14\\

Hypothesis: cluster 23 represents preadipocytes at the start of differentation (the cell states between T3 and T4 in 180831 data that we missed). Cluster 5 represents even more mature metabolic cells and cluster 11 represents more mature ECM cells.\\

Clusters 22 shares most genes with the P branch. Cluster 23 most with the U branch. (see datatable above). These could also represent the preadipocytes at start of differentiation or the cells that transfer back to progenitor cells. \\

```{r}
UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
```

```{r}
#Idents(wolfrum) <- wolfrum@meta.data$seurat_clusters
#preadipocyte_subset <- subset(wolfrum, idents=c(5, 14, 23, 11, 10, 21, 22, 24))
#preadipocyte_subset <- FindVariableFeatures(preadipocyte_subset)
#preadipocyte_subset <- ScaleData(preadipocyte_subset)
#preadipocyte_subset <- RunPCA(object=preadipocyte_subset, npcs=30)
#ElbowPlot(preadipocyte_subset, ndims=30)
#preadipocyte_subset <- FindNeighbors(object = preadipocyte_subset, dims=1:13)
#preadipocyte_subset <- FindClusters(object = preadipocyte_subset, resolution=0.8)
#preadipocyte_subset <- RunTSNE(object = preadipocyte_subset, dims=1:13)
#preadipocyte_subset <- RunUMAP(object = preadipocyte_subset, dims=1:13)
#saveRDS(preadipocyte_subset, '/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.rds')
preadipocyte_subset <- readRDS('output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.rds')

plot_grid(
  UMAPPlot(preadipocyte_subset, group.by='all_data_seurat_clusters', label=T),
  TSNEPlot(preadipocyte_subset, group.by='all_data_seurat_clusters', label=T)
)
```

#Seurat integration of Wolfrum and 10x-180831 data

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
#anchors <- FindIntegrationAnchors(object.list = list(wolfrum, data_180831), dims = 1:20)
#integrated <- IntegrateData(anchorset = anchors, dims = 1:20)
#saveRDS(integrated, '/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.180831.integrated.rds')
integrated <- readRDS('output/seurat_objects/wolfrum/wolfrum.180831.integrated.rds')
integrated@meta.data['dataset'] <- '10x-180831'
integrated@meta.data[which(is.na(integrated@meta.data$branch)), 'dataset'] <- 'Wolfrum'

plot_grid(
  UMAPPlot(integrated, group.by='dataset'),
  UMAPPlot(integrated, group.by='seurat_clusters', label=T),
  UMAPPlot(integrated, group.by='branch'), ncol=2
)
```

These results also confirm that the L branch is closest to cluster 11 and U is closest to the U branch.\\

#Seurat integration of Wolfrum preadiopcyte subset and 10x-180831 data


```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
integrated.subset <- readRDS('output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.180831.integrated.rds')

integrated.subset@meta.data['dataset'] <- '10x-180831'
integrated.subset@meta.data[which(is.na(integrated.subset@meta.data$branch)), 'dataset'] <- 'Wolfrum'
integrated.subset <- AddMetaData(integrated.subset, metadata=wolfrum@meta.data['seurat_clusters'])

plot_grid(
  UMAPPlot(integrated.subset, group.by='dataset'),
  UMAPPlot(integrated.subset, group.by='seurat_clusters', label=T),
  UMAPPlot(integrated.subset, group.by='branch'),
  UMAPPlot(integrated.subset, group.by='timepoint'), ncol=2
)
```

#Predict cell types with Seurat's TransferData

```{r}
wolfrum.predicted_labels <- readRDS('output/seurat_objects/wolfrum/wolfrum.predicted_labels_180831.rds')
```

Used pca, pca.project and cca as dimred for FindTransferAnchors and IntegrateData. \\

Scores for P cells

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca.prediction.score.Progenitor'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca_project.prediction.score.Progenitor'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_cca..prediction.score.Progenitor'), ncol=2
)
```

Scores for U cells

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca.prediction.score.Metabolic'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca_project.prediction.score.Metabolic'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_cca..prediction.score.Metabolic'), ncol=2
)
```

Scores for L cells

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca.prediction.score.ECM'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca_project.prediction.score.ECM'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_cca..prediction.score.ECM'), ncol=2
)
```

For all predictions, change predicted id to NA if max score is below a certain threshold. 

```{r}
assign_labels <- function(colname, threshold=0.5){
  pred_ids <- unlist(as.vector(apply(wolfrum.predicted_labels@meta.data[,c(paste(colname,'.prediction.score.max', sep=''), paste(colname, '.predicted.id', sep=''))], 1, function(x){
    if (x[[1]] < threshold){
      return(NA)
    } else{
      return(x[[2]])
    }
  })))
  return(pred_ids)
}

for (col in c('predictions_pca_project', 'predictions_pca', 'predictions_cca.')){
  for (t in c(0.5, 0.7, 0.9, 0.95, 0.99)){
    preds <- assign_labels(col, t)
    wolfrum.predicted_labels <- AddMetaData(wolfrum.predicted_labels, preds, col.name=paste(col, 'predicted_label', t, sep='.'))
  }
}
```

Threshold for prediction = 0.5

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca.predicted_label.0.5'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca_project.predicted_label.0.5'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_cca..predicted_label.0.5'), ncol=2
)
```

Threshold for prediction = 0.7

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca.predicted_label.0.7'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca_project.predicted_label.0.7'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_cca..predicted_label.0.7'), ncol=2
)
```

Threshold for prediction = 0.9

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca.predicted_label.0.9'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca_project.predicted_label.0.9'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_cca..predicted_label.0.9'), ncol=2
)
```

Only show preadipocytes

```{r}
clusters <- c(24, 22, 11, 10, 23, 5, 14)

adipocytes <- unlist(lapply(wolfrum.predicted_labels@meta.data$RNA_snn_res.0.8, function(x){
  if (x %in% clusters){
    return('preadipocyte')
  } else {
    return('non-preadipocyte')
  }
}))

wolfrum.predicted_labels@meta.data['preadipocytes'] <- adipocytes

for (col in c('predictions_pca_project', 'predictions_pca', 'predictions_cca.')){
  for (t in c(0.5, 0.7, 0.9, 0.95, 0.99)){
    preds <- assign_labels(col, t)
    wolfrum.predicted_labels <- AddMetaData(wolfrum.predicted_labels, preds, col.name=paste(col, 'predicted_label', t, sep='.'))
  }
}

```


#Figures

```{r, fig.height = 4, fig.width = 6, fig.align = "center"}

labels <- unlist(lapply(wolfrum.predicted_labels@meta.data$predictions_pca.predicted_label.0.7, function(x){
  if (is.na(x)){
    return('Non-matching cells')
  } else if(x == 'ECM'){
    return('L')
  } else if(x == 'Metabolic'){
    return('U')
  } else if (x == 'Progenitor'){
    return('P')
  }
}))

wolfrum.predicted_labels@meta.data['predictions_pca.predicted_label.0.7_labels'] <- labels

wolfrum.predicted_labels@meta.data['predictions_pca.predicted_label.0.7_labels'] <- factor(wolfrum.predicted_labels@meta.data$predictions_pca.predicted_label.0.7_labels, levels = c("P", "L", "U", 'Non-matching cells'))
```

```{r, fig.height = 5, fig.width = 8, fig.align = "center"}
colormap.branches <- c(
  P="#ecdd83",
  U="#e27268",
  L="#93c8bc",
  'Non-matching cells'='#7a7a7a')

p_predictions <- UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca.predicted_label.0.7_labels', cols=colormap.branches, no.axes=T) + theme(legend.text=element_text(size=12), legend.key.height=unit(0.4, 'cm'), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm")) + labs(color='Seurat prediction') 
p_predictions
```

```{r}
#save_plot("figures/figures_paper/main_figures/Figure_wolfrum/UMAP_wolfrum_predicted-labels_180831.pdf", p_predictions, base_width=8, base_height=5)
```

```{r, fig.height = 5, fig.width = 6.5, fig.align = "center"}
p_clusters <- UMAPPlot(wolfrum.predicted_labels, group.by='RNA_snn_res.0.8', label=T, no.axes=T, no.legend=T) + theme(legend.position = "none", axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))
p_clusters
```

```{r}
#save_plot("figures/figures_paper/main_figures/Figure_wolfrum/UMAP_wolfrum_clusters.pdf", p_clusters, base_width=6, base_height=5)
```


```{r}
#featureplots
#widht=13
#ucp2_fp <- featureplots_leg$UCP2 + scale_color_gradient(name='Expression', low='gray', #high='blue', guide='colorbar', limits=c(0,5)) + theme(plot.title=element_blank(), #legend.title=element_text(size=20), legend.text=element_text(size=20), legend.key.height = #unit(1.3, 'cm'))


#dcn_fp <- featureplots_noleg$DCN + theme(plot.title=element_blank())

adipoq <- FeaturePlot(wolfrum.predicted_labels, features='ADIPOQ') + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

lipe <- FeaturePlot(wolfrum.predicted_labels, features='LIPE') + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

apod <- FeaturePlot(wolfrum.predicted_labels, features='APOD') + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

dcn <- FeaturePlot(wolfrum.predicted_labels, features='DCN') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5)) 

```

```{r, fig.height = 4, fig.width = 16, fig.align = "center"}
g <- plot_grid(
  adipoq, lipe, apod, dcn, ncol=4,  rel_widths=c(1, 1, 1, 1.3)
)
g
```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/featureplots.pdf", g, base_width=16, base_height=4)
```

EBF2 and LEP

```{r}
ebf2 <- FeaturePlot(wolfrum.predicted_labels, features='EBF2') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

pparg <- FeaturePlot(wolfrum.predicted_labels, features='PPARG') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

lep <- FeaturePlot(wolfrum.predicted_labels, features='LEP') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

ucp1 <- FeaturePlot(wolfrum.predicted_labels, features='UCP1') + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=20), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar') 

ebf2vln <- VlnPlot(wolfrum.predicted_labels, features='EBF2', group.by='RNA_snn_res.0.8', pt.size=-1) + NoLegend()

lepvln <- VlnPlot(wolfrum.predicted_labels, features='LEP', group.by='RNA_snn_res.0.8', pt.size=0.1) + NoLegend()

ucp1vln <- VlnPlot(wolfrum.predicted_labels, features='UCP1', group.by='RNA_snn_res.0.8', pt.size=0.1) + NoLegend()

ppargvln <- VlnPlot(wolfrum.predicted_labels, features='PPARG', group.by='RNA_snn_res.0.8', pt.size=0.1) + NoLegend()

```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ebf2.pdf", ebf2, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ebf2vln.pdf", ebf2vln, base_width=6, base_height=2)

#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/pparg.pdf", pparg, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ppargvln.pdf", ppargvln, base_width=6, base_height=2)

#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ucp1.pdf", ucp1, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/ucp1vln.pdf", ucp1vln, base_width=6, base_height=2)

#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/lep.pdf", lep, base_width=5, base_height=4)
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/lepvln.pdf", lepvln, base_width=5, base_height=2)

```


```{r}
UMAPPlot(wolfrum.predicted_labels, label=T, group.by='RNA_snn_res.0.8') 
```

```{r, fig.height = 6, fig.width = 12, fig.align = "center"}
VlnPlot(wolfrum.predicted_labels, features=c('EBF2', 'PPARG'), group.by='RNA_snn_res.0.8', pt.size=-1, ncol=1)
```


```{r, fig.height = 6, fig.width = 6, fig.align = "center"}
#EBF2 and PPARG
#PPARG: 5, 14, 23
#EBF2: 10, 11, 21, 23

new_labels <- unlist(lapply(wolfrum.predicted_labels@meta.data$RNA_snn_res.0.8, function(x){
  if (x %in% c(5, 14, 23, 10, 11, 21, 23)){
    return(x)
  } else {
    return(NA)
  }
}))

wolfrum.predicted_labels@meta.data['labels_clusters_preadipocytes'] <- new_labels
p <- UMAPPlot(wolfrum.predicted_labels, group.by='labels_clusters_preadipocytes', label=T) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm")) + NoLegend()
p
```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/UMAP_adipocyte_clusters2.pdf", p, base_width=6, base_height=5)
```


```{r, fig.height = 4, fig.width = 8.5, fig.align = "center"}
ebf2 <- FeaturePlot(wolfrum.predicted_labels, features='EBF2', pt.size=0.5) + NoLegend() + NoAxes() + theme(plot.title = element_text(size=20)) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5))

pparg <- FeaturePlot(wolfrum.predicted_labels, features='PPARG', pt.size=0.5) + NoAxes() + theme(plot.title = element_text(size=20), legend.text=element_text(size=15), legend.key.height = unit(1.3, 'cm')) + scale_color_gradient(name='Expression', low='gray', high='blue', guide='colorbar', limits=c(0,5)) 

g <- plot_grid(
  ebf2, pparg, ncol=2, rel_widths = c(1, 1.3)
)

g
```

```{r}
#save_plot("../figures/figures_paper/main_figures/Figure_wolfrum/EBF2_PPARG_ptsize0.5.pdf", g, base_width=8.5, base_height=4)
```


```{r, fig.height = 10, fig.width = 8, fig.align = "center"}
plots <- VlnPlot(wolfrum.predicted_labels, features=c('ADIPOQ', 'LIPE', 'PLIN4', 'FABP4', 'ADIRF', 'APOD', 'MGP', 'DCN', 'CCDC80', 'PLAC9'), group.by='RNA_snn_res.0.8', pt.size=-1, combine=F)

for (i in 1:length(plots)){
  if (i == length(plots)){
    plots[[i]] <- plots[[i]] + NoLegend() + 
      theme(plot.title=element_blank(), 
            axis.title.y=element_blank(), 
            axis.line.x=element_blank(),
            axis.text.x=element_text(angle=0, size=12),
            plot.margin = unit(c(0, 0, 0, 0), "cm")) + 
      labs(x='Cluster')
  } else {
    plots[[i]] <- plots[[i]] + NoLegend() + 
      theme(plot.title=element_blank(), 
            axis.title.y=element_blank(), 
            axis.line.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_blank(),
            axis.title.x=element_blank(),
            plot.margin = unit(c(0, 0, 0, 0), "cm"))
  }
}

vlnplts <- plot_grid(plotlist=plots, ncol=1, rel_heights=c(1,1,1,1,1,1,1,1,1,1.6))
vlnplts
```


```{r}
#save_plot("figures/figures_paper/main_figures/Figure_wolfrum/violinplots.pdf", p, base_width=6, base_height=8)
```

Supplementary figure

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
integrated <- readRDS('output/seurat_objects/wolfrum/wolfrum.180831.integrated.rds')
sfig <- plot_grid(
  UMAPPlot(integrated, group.by='dataset'),
  UMAPPlot(integrated, group.by='State.labels', cols=colormap.branches),
  UMAPPlot(integrated, group.by='RNA_snn_res.0.8', label=T) + NoLegend(), ncol=2
)
sfig
```

```{r}
#save_plot("figures/figures_paper/supplementary_figures/wolfrum/integration.wolfrum.180831.pdf", sfig, base_width=12, base_height=8)
```

