---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: 
      smooth_scroll: true
---

```{r}
library(Seurat)
library(scmap)
library(SingleCellExperiment)
library(dplyr)
```

Data loading and inspection of the metadata.

```{r}
load('/data/pub-others/tabula_muris/figshare/180126-facs/maca.seurat_obj.facs.figshare_180126.RData')
head(seurat_obj@meta.data)
UpdateSeuratObject(seurat_obj)
```


```{r}
sce_maca <- as.SingleCellExperiment(seurat_obj)
wolfrum <- readRDS('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.compute.seurat_obj.rds')
sce_wolfrum <- as.SingleCellExperiment(wolfrum)

#convert maca gene names to uppercase to match 10x gene names
rowData(sce_maca)['feature_symbol'] <- unlist(lapply(rowData(sce_maca)$gene, function(x){return(toupper(x))}))
rowData(sce_10x)['feature_symbol'] <- rowData(sce_10x)$gene

counts(sce_10x) <- as.matrix(counts(sce_10x))
logcounts(sce_10x) <- as.matrix(logcounts(sce_10x))

counts(sce_maca) <- as.matrix(counts(sce_maca))
logcounts(sce_maca) <- as.matrix(logcounts(sce_maca))
sce_maca <- selectFeatures(sce_maca, suppress_plot = FALSE)
```

Celltypes in the fat dataset

```{r}
seurat_obj@meta.data %>% filter(tissue=="Fat") %>% distinct(tissue_cell_type)
```

Subsetting and preparing the data.

```{r}
maca_fat <- SubsetData(SetAllIdent(seurat_obj, id='tissue'), ident.use="Fat")
sce_maca_fat <- as.SingleCellExperiment(maca_fat)
rowData(sce_maca_fat)['feature_symbol'] <- unlist(lapply(rowData(sce_maca_fat)$gene, function(x){return(toupper(x))}))
counts(sce_maca_fat) <- as.matrix(counts(sce_maca_fat))
logcounts(sce_maca_fat) <- as.matrix(logcounts(sce_maca_fat))
sce_maca_fat <- selectFeatures(sce_maca_fat, suppress_plot = FALSE)
```

Setting the right column for clustering.

```{r}
sce_maca_fat <- indexCluster(sce_maca_fat, cluster_col = 'cell_ontology_class')
```

Predicting cell types in our dataset.

```{r}
scmapCluster_results_fat <- scmapCluster(
  projection = sce_10x, 
  index_list = list(
    sce_maca_fat = metadata(sce_maca_fat)$scmap_cluster_index
  ),
  threshold=0.5  #default=0.7 
)
```

Number of predictions for each annotation for the whole dataset and for the mixture cluster.

```{r}
pred_fat <- as.data.frame(table(scmapCluster_results_fat$scmap_cluster_labs))
pred_fat <- pred_fat[order(-pred_fat$Freq),]

pred_mixt_fat <- as.data.frame(table(scmapCluster_results_fat$scmap_cluster_labs[which(colData(sce_10x)$res.0.5 %in% 12), 'sce_maca_fat']))

scmap_nr_predictions <- merge(pred_fat, pred_mixt_fat, by='Var1', suffixes=c('.total', '.mixture'))
scmap_nr_predictions
```

Interestingly, a lot of epithelial cell predictions in the mixture cluster and not that much mesenchymal stem cell predictions.

```{r}
predicted_labels_fat <- as.data.frame(
    row.names=rownames(sce_10x@colData), 
    x=as.vector(scmapCluster_results_fat$scmap_cluster_labs))
names(predicted_labels_fat) <- 'predicted_labels_fat'
all10x <- AddMetaData(all10x, metadata=predicted_labels_fat, col.name='predicted_labels_fat')
t1 <- TSNEPlot(all10x, group.by='predicted_labels_fat', pt.size=0.1)
save_plot("/projects/pytrik/sc_adipose/analyze_10x_fluidigm/data/plots_slides/scmap.pdf", t1, base_width=8, base_height = 5)
```

```{r}
TSNEPlot(all10x, group.by='sample_name', pt.size=0.1, do.label=T)
```


```{r}
#saveRDS(all10x, 'output/10x-180504')
```

```{r}
write.table(scmap_nr_predictions, 'tables/10x-180504-scmap-numbers')
```

