---
title: "R Notebook"
output: html_notebook
---


```{r message=F}
library(Seurat)
library(monocle)
library(cowplot)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(DT)
```

```{r}
wolfrum <- readRDS('/projects/timshel/sc-scheele_lab_adipose_fluidigm_c1/data-wolfrum/wolfrum.compute.seurat_obj.rds')
data_180831 <- readRDS('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/180831/10x-180831-S3')
```



##Which clusters are adipocytes/preadipocytes?

```{r fig.height = 4, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum, group.by='orig.ident', label=T),
  UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
)
```

```{r}
markers <- read.table('../output/markergenes/wolfrum/markers_wolfrum.compute.seurat_obj.rds_seurat_clusters_negbinom', sep='\t', header=T)
```

Top 10 positive markers per cluster 

```{r}
pos_markers_top10 <- markers %>% 
  group_by(cluster) %>% 
  top_n(n=10, wt=avg_logFC)
  
neg_markers_top20 <- markers %>% 
  group_by(cluster) %>% 
  top_n(n=6, wt=desc(avg_logFC))

pos_markers_top10
```

Which clusters are the preadipocytes? Check by plotting some of the DE genes between T1T2T3 and T4T5 in the 180831 dataset.

```{r}
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_T1T2T3_T4T5 <- read.table('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/markergenes/180831/markers_10x-180831_time_combined_negbinom', header=T)
markers_T1T2T3_T4T5 <- markers_T1T2T3_T4T5[order(-markers_T1T2T3_T4T5$avg_logFC),]
markers_T1T2T3 <- markers_T1T2T3_T4T5[which(markers_T1T2T3_T4T5$cluster == 1),]
markers_T4T5 <- markers_T1T2T3_T4T5[which(markers_T1T2T3_T4T5$cluster == 2),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_T1T2T3$gene)[1:10], as.vector(markers_T4T5$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

How are they expressed in the Wolfrum data?

```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(wolfrum, features=c(as.vector(markers_T1T2T3$gene)[1:10], as.vector(markers_T4T5$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

Not all genes are expressed clearly in specific clusters. It looks like cluster 22, 21, 5, 14, 23, 11 and 10 are the preadipocytes. 

```{r}
UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
```

Also plot the top 10 ECM and Metabolic markers.

```{r}
#DE genes between T1T2T3 and T4T5 in the 10x-180831 data.
markers_u_l <- read.table('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/markergenes/180831/markers_10x-180831_upperbranch_lowerbranch_negbinom', sep='\t', header=T)
markers_u <- markers_u_l[order(-markers_u_l$avg_logFC),]
markers_l <- markers_u_l[order(markers_u_l$avg_logFC),]
```

How do these genes look in the 180831 data?

```{r, fig.height = 12, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(data_180831, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=4)
```

And how are they expressed in the Wolfrum data?

```{r, fig.height = 50, fig.width = 12, fig.align = "center"}
plots <- FeaturePlot(wolfrum, features=c(as.vector(markers_u$gene)[1:10], as.vector(markers_l$gene)[1:10]), pt.size=1, combine=F)
plot_grid(plotlist=plots, ncol=2)
```

The U and L branch markers are clearly expressed in the two clusters in the top right of the UMAP plot. 

Check how many of the marker genes are found in each cluster. 

```{r}
get_gene_overlap_per_cluster <- function(min_logFC, n_genes){
  pos_markers <- markers[markers$avg_logFC > min_logFC,]

  genes_P <- c(as.vector(markers_T1T2T3$gene)[1:n_genes])
  #genes_P <- c(as.vector(markers_T1T2T3$gene)[1:n_genes], as.vector(markers_T4T5$gene)[1:n_genes])
  genes_U <- as.vector(markers_u$gene)[1:n_genes]
  genes_L <- as.vector(markers_l$gene)[1:n_genes]
  all_genes <- unique(c(genes_P, genes_U, genes_L))
  
  df <- as.data.frame(matrix(ncol=4, nrow=length(unique(pos_markers$cluster))))
  colnames(df) <- c('total_overlap', 'overlap_P', 'overlap_L', 'overlap_U')
  rownames(df) <- unique(pos_markers$cluster)

  for (i in 1:length(unique(pos_markers$cluster))){
    genes_cluster <- as.vector(pos_markers[pos_markers$cluster == i,]$gene)
    df[i, 'overlap_P'] <- round(length(intersect(genes_cluster, genes_P)) / length(genes_P), 2)
    df[i, 'overlap_L'] <- round(length(intersect(genes_cluster, genes_L)) / length(genes_L), 2)
    df[i, 'overlap_U'] <- round(length(intersect(genes_cluster, genes_U)) / length(genes_U), 2)
    df[i, 'total_overlap'] <- round(length(intersect(genes_cluster, all_genes)) / length(all_genes), 2)
  }
  return(df)
}

print_top_clusters <- function(df){
  print(paste('P: ', toString(rownames(df[order(-df$overlap_P),][1:3,])), sep=''))
  print(paste('L: ', toString(rownames(df[order(-df$overlap_L),][1:3,])), sep=''))
  print(paste('U: ', toString(rownames(df[order(-df$overlap_U),][1:3,])), sep=''))
}

```

Table shows the percentage of genes with avgLogFC > 0.7 that was found in the cluster. Sort on columns to get the top clusters.

```{r}
logfc_0.5_genes_100 <- get_gene_overlap_per_cluster(0.5, 100)
datatable(logfc_0.5_genes_100)
```

Check the top 5 clusters per branch for different n_genes and min logFC

```{r}
logfc <- c(0.25, 0.5, 0.7)
n_genes <- c(100, 100, 100)
for (i in 1:length(logfc)){
  print(paste('min_logFC: ', logfc[i], ' | n_genes: ', n_genes[i], sep=''))
  df <- get_gene_overlap_per_cluster(logfc[i], n_genes[i])
  print_top_clusters(df)
}
```

There is some overlap between branches which is good. Cluster 11 is shared between P and L, cluster 14 is shared between P and U and cluster 23 is shared between L and U. This would indicate that cluster 24 contains most immature preadipocytes, cluster 10 contains most mature L branch cells and cluster 22 contains most mature U branch cells.\\

Based on the results:\\
P = 24\\
L = 11\\
U = 14\\

Hypothesis: cluster 23 represents preadipocytes at the start of differentation (the cell states between T3 and T4 in 180831 data that we missed). Cluster 5 represents even more mature metabolic cells and cluster 11 represents more mature ECM cells.\\

Clusters 22 shares most genes with the P branch. Cluster 23 most with the U branch. (see datatable above). These could also represent the preadipocytes at start of differentiation or the cells that transfer back to progenitor cells. \\

```{r}
UMAPPlot(wolfrum, group.by='seurat_clusters', label=T)
```

```{r}
#Idents(wolfrum) <- wolfrum@meta.data$seurat_clusters
#preadipocyte_subset <- subset(wolfrum, idents=c(5, 14, 23, 11, 10, 21, 22, 24))
#preadipocyte_subset <- FindVariableFeatures(preadipocyte_subset)
#preadipocyte_subset <- ScaleData(preadipocyte_subset)
#preadipocyte_subset <- RunPCA(object=preadipocyte_subset, npcs=30)
#ElbowPlot(preadipocyte_subset, ndims=30)
#preadipocyte_subset <- FindNeighbors(object = preadipocyte_subset, dims=1:13)
#preadipocyte_subset <- FindClusters(object = preadipocyte_subset, resolution=0.8)
#preadipocyte_subset <- RunTSNE(object = preadipocyte_subset, dims=1:13)
#preadipocyte_subset <- RunUMAP(object = preadipocyte_subset, dims=1:13)
#saveRDS(preadipocyte_subset, '/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.rds')
preadipocyte_subset <- readRDS('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.rds')

plot_grid(
  UMAPPlot(preadipocyte_subset, group.by='all_data_seurat_clusters', label=T),
  TSNEPlot(preadipocyte_subset, group.by='all_data_seurat_clusters', label=T)
)
```

##Seurat integration of Wolfrum and 10x-180831 data

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
#anchors <- FindIntegrationAnchors(object.list = list(wolfrum, data_180831), dims = 1:20)
#integrated <- IntegrateData(anchorset = anchors, dims = 1:20)
#saveRDS(integrated, '/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.180831.integrated.rds')
integrated <- readRDS('/projects/pytrik/sc_adipose/analyze_10x_fluidigm/10x-adipocyte-analysis/output/seurat_objects/wolfrum/wolfrum.180831.integrated.rds')
integrated@meta.data['dataset'] <- '10x-180831'
integrated@meta.data[which(is.na(integrated@meta.data$branch)), 'dataset'] <- 'Wolfrum'

plot_grid(
  UMAPPlot(integrated, group.by='dataset'),
  UMAPPlot(integrated, group.by='seurat_clusters', label=T),
  UMAPPlot(integrated, group.by='branch'), ncol=2
)
```

These results also confirm that the L branch is closest to cluster 11 and U is closest to the U branch.\\

##Seurat integration of Wolfrum preadiopcyte subset and 10x-180831 data


```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
integrated.subset <- readRDS('output/seurat_objects/wolfrum/wolfrum.preadipocyte_subset.180831.integrated.rds')

integrated.subset@meta.data['dataset'] <- '10x-180831'
integrated.subset@meta.data[which(is.na(integrated.subset@meta.data$branch)), 'dataset'] <- 'Wolfrum'
integrated.subset <- AddMetaData(integrated.subset, metadata=wolfrum@meta.data['seurat_clusters'])

plot_grid(
  UMAPPlot(integrated.subset, group.by='dataset'),
  UMAPPlot(integrated.subset, group.by='seurat_clusters', label=T),
  UMAPPlot(integrated.subset, group.by='branch'),
  UMAPPlot(integrated.subset, group.by='timepoint'), ncol=2
)
```

##Predict cell types with Seurat's TransferData

```{r}
wolfrum.predicted_labels <- readRDS('output/seurat_objects/wolfrum/wolfrum.predicted_labels_180831.rds')
```

Used pca, pca.project and cca as dimred for FindIntegrationAnchors and IntegrateData. \\

Scores for P cells

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca.prediction.score.Progenitor'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca_project.prediction.score.Progenitor'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_cca..prediction.score.Progenitor'), ncol=2
)
```

Scores for U cells

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca.prediction.score.Metabolic'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca_project.prediction.score.Metabolic'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_cca..prediction.score.Metabolic'), ncol=2
)
```

Scores for L cells

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca.prediction.score.ECM'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_pca_project.prediction.score.ECM'),
  FeaturePlot(wolfrum.predicted_labels, features='predictions_cca..prediction.score.ECM'), ncol=2
)
```

For all predictions, change predicted id to NA if max score is below a certain threshold. 

```{r}
assign_labels <- function(colname, threshold=0.5){
  pred_ids <- unlist(as.vector(apply(wolfrum.predicted_labels@meta.data[,c(paste(colname,'.prediction.score.max', sep=''), paste(colname, '.predicted.id', sep=''))], 1, function(x){
    if (x[[1]] < threshold){
      return(NA)
    } else{
      return(x[[2]])
    }
  })))
  return(pred_ids)
}

for (col in c('predictions_pca_project', 'predictions_pca', 'predictions_cca.')){
  for (t in c(0.5, 0.7, 0.9, 0.95, 0.99)){
    preds <- assign_labels(col, t)
    wolfrum.predicted_labels <- AddMetaData(wolfrum.predicted_labels, preds, col.name=paste(col, 'predicted_label', t, sep='.'))
  }
}
```

Threshold for prediction = 0.7

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca.predicted_label.0.7'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca_project.predicted_label.0.7'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_cca..predicted_label.0.7'), ncol=2
)
```

Threshold for prediction = 0.9

```{r, fig.height = 8, fig.width = 12, fig.align = "center"}
plot_grid(
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca.predicted_label.0.'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_pca_project.predicted_label.0.9'),
  UMAPPlot(wolfrum.predicted_labels, group.by='predictions_cca..predicted_label.0.9'), ncol=2
)
```


##Monocle

Ran Monocle on the Wolfrum subset with the same gene list that was used to build the. 

```{r}
monocle_results <- readRDS('output/monocle/wolfrum/wolfrum.preadipocyte_subset.rds-monocle-monocle_genelist_T1T2T3_T4T5_res.1.5')

predictions <- wolfrum.predicted_labels@meta.data['predictions_pca.predicted.id']
sorted_predictions <- predictions[rownames(pData(monocle_results)),]
pData(monocle_results)['predicted_labels'] <- sorted_predictions

```


```{r, fig.height = 4, fig.width = 12, fig.align = "center"}
plot_grid(
  plot_cell_trajectory(monocle_results, color_by='State'),
  plot_cell_trajectory(monocle_results, color_by='predicted_labels')
)
```

```{r}
Idents(wolfrum.predicted_labels) <- wolfrum.predicted_labels$seurat_clusters
preadipocyte_subset <- subset(wolfrum.predicted_labels, idents=c(5, 14, 23, 11, 10, 21, 22, 24))
preadipocyte_subset <- AddMetaData(preadipocyte_subset, pData(monocle_results)['State'])
```


```{r}
get_ratios <- function(data, col1, col2){
  states <- unique(data@meta.data[,col1])
  values <- unique(data@meta.data[,col2])
  df <- as.data.frame(matrix(ncol=length(values)+1, nrow=length(states)))
  colnames(df) <- c('n', values)
  rownames(df) <- states
  for (state in states){
    n_state = length(which(data@meta.data[col1] == state))
    df[state, 'n'] <- n_state
    for (value in values){
    	n_state_value <- length(which(data@meta.data[col1] == state & data@meta.data[col2] == value))
	    perc_state_value <- n_state_value / n_state
	    df[state, value] <- round(perc_state_value, 2)
    }
  }
  return(df)
}
```

```{r}
states.prediction <- get_ratios(data=preadipocyte_subset, col1='State', col2='predictions_pca.predicted.id')
states.prediction
```

```{r}
states.prediction <- get_ratios(data=preadipocyte_subset, col2='State', col1='predictions_pca.predicted.id')
states.prediction
```




```{r}
get_gene_overlap_per_cluster <- function(min_logFC, n_genes){
  pos_markers <- markers[markers$avg_logFC > min_logFC,]

  all_genes <- mixture_markers$gene[1:n_genes]
  
  df <- as.data.frame(matrix(ncol=4, nrow=length(unique(pos_markers$cluster))))
  colnames(df) <- c('total_overlap', 'overlap_P', 'overlap_L', 'overlap_U')
  rownames(df) <- unique(pos_markers$cluster)

  for (i in 1:length(unique(pos_markers$cluster))){
    genes_cluster <- as.vector(pos_markers[pos_markers$cluster == i,]$gene)
    df[i, 'overlap_P'] <- round(length(intersect(genes_cluster, genes_P)) / length(genes_P), 2)
    df[i, 'overlap_L'] <- round(length(intersect(genes_cluster, genes_L)) / length(genes_L), 2)
    df[i, 'overlap_U'] <- round(length(intersect(genes_cluster, genes_U)) / length(genes_U), 2)
    df[i, 'total_overlap'] <- round(length(intersect(genes_cluster, all_genes)) / length(all_genes), 2)
  }
  return(df)
}

print_top_clusters <- function(df){
  print(paste('P: ', toString(rownames(df[order(-df$overlap_P),][1:3,])), sep=''))
  print(paste('L: ', toString(rownames(df[order(-df$overlap_L),][1:3,])), sep=''))
  print(paste('U: ', toString(rownames(df[order(-df$overlap_U),][1:3,])), sep=''))
}
```

